### 2017-04-20 paper/main.tex
paper initial

\documentclass{article}

\usepackage{fullpage}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage[colorlinks,linkcolor=dblue,filecolor=black,citecolor=dblue,urlcolor=dblue]{hyperref}
\usepackage{etex}
\bibliographystyle{alpha}
\usepackage[english]{babel}
\usepackage{amsfonts,amssymb,amsmath,sectsty,url}
\usepackage{mathrsfs}

\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{color}
%\usepackage{mathtools}
\usepackage[all]{xy}
\usepackage{ctable}
% \usetikzlibrary{calc}

\usepackage{graphicx}
\usepackage{ifpdf}
\usepackage{multirow}
\usepackage{multicol,color}
\usepackage{tablefootnote}

\usepackage{amsthm}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}%[section]
\newtheorem{proposition}[theorem]{Proposition}%[section]

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}%[section]
\newtheorem{construction}[theorem]{Construction}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{fact}[theorem]{Fact}%[section]
\newtheorem{observation}[theorem]{Observation}
\newtheorem{claim}[theorem]{Claim}%[section]

\input{symbols}

\newcommand{\mr}[1]{{\textcolor{lblue}{\textbf{Mariana:} #1}}}
\newcommand{\vc}[1]{{\textcolor{lred}{\textbf{Valerie:} #1}}}
\newcommand{\vp}[1]{{\textcolor{dblue}{\textbf{Valerio:} #1}}}

\title{Multiparty Linear Regression via SPDZ}
\author{Valerie Chen \and Valerio Pastro \and Mariana Raykova}

\begin{document}

\titlepage
\maketitle

\pagebreak

\section{Preliminaries}
\vp{stuff}

\section{Result}

\end{document}

### 2017-06-01 paper/main.tex
added results

@@
 \usepackage{amssymb}
-
+\usepackage{graphicx}
 \usepackage[colorlinks,linkcolor=dblue,filecolor=black,citecolor=dblue,urlcolor=dblue]{hyperref}
@@
 
+\section{Experiments}
+The experiments were run on the SPDZ framework, using a 128 bit and 256 bit prime field respectively.
+In the 128 bit prime field, the bits were allotted with 28 bits after the decimal place to match the 32 bit example in the original paper and 60 bits after the decimal for the original 64 bit case.
+Compared original paper, we needed more space to the left of the decimal for comparison purposes and operations using the fixed point type.
+However, we believe this is still a fair comparison because we are not working with more precision after the decimal place and if anything using more bits should actually slow down our calculations.
+For each generated matrix of varying dimensions, the features were split evenly between two parties.
+We then wanted to see if linear decomposition algorithms like LDLT, Cholesky, and CGD (with 5, 10, 15, and 20 iterations) were able to get comparable if not better results.
+
+First, we ran the two algorithms two gauge time and error rates for matrices of size 10, 20, 50, 100, 200, and 500, with 200000 features each.
+Secondly, we ran on a size 20 matrices with condition numbers varying from 1 to 10.
+Both parts were completed for both bit sizes to evaluate performance differences.
+To model after the original experiment, we ran iterations of 5, 10, 15, and 20 for the CGD algorithm but added on as an extension of 25 iterations to see if necessary (i.e.
+the errors were not comparable for all sizes or condition numbers), there was still room for improvement.
+
 \section{Result}
 
+\begin{figure}[h]
+  \includegraphics[width=\linewidth]{results.jpg}
+  \caption{(Left) Comparison between different methods of solving linear systems: running time in seconds for Cholesky, LDLT, and CGD (with 5, 10, 15, and 20 iterations) as a function of input dimension.
+(Middle) Accuracy of CGD as a function of the input on the condition number of the input matrix A with d = 20.
+(Right) Accuracy of Cholesky, LDLT, and CGD, as a function of the input dimensionality d.
+(Top) Fixed-point numbers with the b = 80 bits, with 60 in the fractional part.
+(Bottom) b = 41, with 28 in the fractional part}
+  \label{fig:result1}
+\end{figure}
+
+As expected, utilizing more bits gives much better accuracy.
+In both cases, the error results are comparable to the garbled circuits paper.
+However, we did find that the original experiment truncated the generated data at 12 decimal places, however we modified it so that we could work with the full representation of the number to get a better understanding of how accurate our algorithm is.
+
+The run times for SPDZ are always comparable if not much faster.
+In terms of run-time, we observe that both Cholesky and LDLT were faster for smaller dimensions when d $\leq$ 100 and CGD is faster for larger dimensions when d > 100.
+This is consistent with the findings in the original experiment.
+For dimensions 100 and smaller, when Cholesky and LDLT are faster, they comparative differences are not that big (LDLT with 1.6 seconds compared to CGD 20 with 60.89 seconds) when compared to the largest differences between the larger sizes (LDLT with 1747 seconds and CGD5 with 97 seconds).
+
+[insert table comparison here??]
+
+In the condition number experiment, for CGD iterations of 5, 10, and 15, error increases as condition number increases.
+For example, in CGD 10 a small condition number may have an error of 10e-9 yet a larger condition number could have an error as large as 10e-3.
+However, in the CGD 20 case, the condition number does not seem to have as much of effect as the range of errors is much less., with the error varying only form 10e-11 to 10e-14 keeping in mind these are much smaller increments.
+(why??
+Add some explanation about how condition number affects errors in a matrix) For the 60 bit precision case, we chose to run more iterations because in the CGD 20 line, there is a noticeable upward slope in the scatter plot, indicating that still as condition numbers increased, the errors were increasing as well.
+Indeed, 25 iterations showed some improvement as we can see a flatter trend in the points.
+
 \end{document}

### 2017-06-14 paper/main.tex
formatting

@@
   \caption{(Left) Comparison between different methods of solving linear systems: running time in seconds for Cholesky, LDLT, and CGD (with 5, 10, 15, and 20 iterations) as a function of input dimension.
-(Middle) Accuracy of CGD as a function of the input on the condition number of the input matrix A with d = 20.
-(Right) Accuracy of Cholesky, LDLT, and CGD, as a function of the input dimensionality d.
-(Top) Fixed-point numbers with the b = 80 bits, with 60 in the fractional part.
-(Bottom) b = 41, with 28 in the fractional part}
+(Middle) Accuracy of CGD as a function of the input on the condition number of the input matrix $A$ with $d = 20$.
+(Right) Accuracy of Cholesky, LDLT, and CGD, as a function of the input dimensionality $d$.
+(Top) Fixed-point numbers with the $b = 80$ bits, with 60 in the fractional part.
+(Bottom) $b = 41$, with 28 in the fractional part}
   \label{fig:result1}
@@
 The run times for SPDZ are always comparable if not much faster.
-In terms of run-time, we observe that both Cholesky and LDLT were faster for smaller dimensions when d $\leq$ 100 and CGD is faster for larger dimensions when d > 100.
+In terms of run-time, we observe that both Cholesky and LDLT were faster for smaller dimensions when $d \leq 100$ and CGD is faster for larger dimensions when $d > 100$.
 This is consistent with the findings in the original experiment.
@@
 
-[insert table comparison here??]
+\vc{insert table comparison here??}
 
 In the condition number experiment, for CGD iterations of 5, 10, and 15, error increases as condition number increases.
-For example, in CGD 10 a small condition number may have an error of 10e-9 yet a larger condition number could have an error as large as 10e-3.
-However, in the CGD 20 case, the condition number does not seem to have as much of effect as the range of errors is much less., with the error varying only form 10e-11 to 10e-14 keeping in mind these are much smaller increments.
-(why??
-Add some explanation about how condition number affects errors in a matrix) For the 60 bit precision case, we chose to run more iterations because in the CGD 20 line, there is a noticeable upward slope in the scatter plot, indicating that still as condition numbers increased, the errors were increasing as well.
+For example, in CGD 10 a small condition number may have an error of $\mathtt{10e-9}$ yet a larger condition number could have an error as large as $\mathtt{10e-3}$.
+However, in the CGD 20 case, the condition number does not seem to have as much of effect as the range of errors is much less., with the error varying only form $\mathtt{10e-11}$ to $\mathtt{10e-14}$ keeping in mind these are much smaller increments.
+\vc{why??
+Add some explanation about how condition number affects errors in a matrix} For the 60 bit precision case, we chose to run more iterations because in the CGD 20 line, there is a noticeable upward slope in the scatter plot, indicating that still as condition numbers increased, the errors were increasing as well.
 Indeed, 25 iterations showed some improvement as we can see a flatter trend in the points.

### 2017-06-15 paper/main.tex
new charts haven't changed words

@@
 
+To better fine tune the CGD and Cholesky algorithms to SPDZ, we ran a few experiments to test hypotheses for
+
 First, we ran the two algorithms two gauge time and error rates for matrices of size 10, 20, 50, 100, 200, and 500, with 200000 features each.
@@
 \begin{figure}[h]
-  \includegraphics[width=\linewidth]{results.jpg}
+  \includegraphics[width=\linewidth]{sqrt.jpg}
+  \caption{ (Top) 64 bit (Bottom) 32 bit results }
+  \label{fig:result1}
+\end{figure}
+
+Comparing square root for Cholesky.
+In SPDZ, we hypothesized that division would be time consuming, however the behavior does not seem to differ much both in terms of run time and error...
+Babylonian vs Newton method differ in how many divisions are used in the algorithm.
+From the results we can see that using different operations don't change the final result.
+This might be because the way that division is implemented in SPDZ..
+
+For CGD, we decide to test if L2 norm vs Linfinity would make a difference. One utilizes comparisons and the other addition and square root. 
+\begin{figure}[h]
+  \includegraphics[width=\linewidth]{l2.jpg}
+  \caption{ (Top) 64 bit (Bottom) 32 bit results }
+  \label{fig:result1}
+\end{figure}
+
+Averaged results Averaged column 3 over 30 samples with n = 100000 (avg cn 1.77?)
+
+\vc{compare ranges of error between them???}
+
+N = 100000
+\begin{figure}[h]
+  \includegraphics[width=\linewidth]{results1.jpg}
   \caption{(Left) Comparison between different methods of solving linear systems: running time in seconds for Cholesky, LDLT, and CGD (with 5, 10, 15, and 20 iterations) as a function of input dimension.
@@
 (Bottom) $b = 41$, with 28 in the fractional part}
-  \label{fig:result1}
+  \label{fig:result2}
 \end{figure}
+
+N = 1000
+\begin{figure}[h]
+  \includegraphics[width=\linewidth]{results2.jpg}
+  \caption{(Left) Comparison between different methods of solving linear systems: running time in seconds for Cholesky, LDLT, and CGD (with 5, 10, 15, and 20 iterations) as a function of input dimension.
+(Middle) Accuracy of CGD as a function of the input on the condition number of the input matrix $A$ with $d = 20$.
+(Right) Accuracy of Cholesky, LDLT, and CGD, as a function of the input dimensionality $d$.
+(Top) Fixed-point numbers with the $b = 80$ bits, with 60 in the fractional part.
+(Bottom) $b = 41$, with 28 in the fractional part}
+  \label{fig:result2}
+\end{figure}
+
+\vc{talk about outliers in these cases, skews results}
 
@@
 For dimensions 100 and smaller, when Cholesky and LDLT are faster, they comparative differences are not that big (LDLT with 1.6 seconds compared to CGD 20 with 60.89 seconds) when compared to the largest differences between the larger sizes (LDLT with 1747 seconds and CGD5 with 97 seconds).
+It is also important to note that the time differences between 32 and 64 bits are not that large- only blank percent more compared to the garbled circuits result where doubling the number of bits roughly increased time by an entire magnitude of 10.
+
+Condition number plot
+\begin{figure}[h]
+  \includegraphics[width=\linewidth]{conditionnumber.jpg}
+  \caption{(Left) Comparison between different methods of solving linear systems: running time in seconds for Cholesky, LDLT, and CGD (with 5, 10, 15, and 20 iterations) as a function of input dimension.
+(Middle) Accuracy of CGD as a function of the input on the condition number of the input matrix $A$ with $d = 20$.
+(Right) Accuracy of Cholesky, LDLT, and CGD, as a function of the input dimensionality $d$.
+(Top) Fixed-point numbers with the $b = 80$ bits, with 60 in the fractional part.
+(Bottom) $b = 41$, with 28 in the fractional part}
+  \label{fig:result3}
+\end{figure}
 
@@
 
+- increasing d and increasing the condition number will result in higher errors, so you need more iterations of CGD to get comparable results.
+- for high condition numbers, Cholesky outperforms CGD (i.e., gets more accurate results with similar running time).
+- for high values of d, but low condition numbers, CGD outperforms Cholesky in both running time and accuracy (that's basically the right column in Figure 6 in the PETS paper)
+Regarding the correlation between n, d, and the condition number, increasing d and keeping n constant will result in a higher expected condition number.
+(confirmed with Phillipp)
+
 \end{document}

### 2017-08-01 paper/main.tex
new charts
Add files via upload

@@
   \includegraphics[width=\linewidth]{sqrt.jpg}
-  \caption{ (Top) 64 bit (Bottom) 32 bit results }
+  \caption{ (Top) 64 bit (Bottom) 32 bit results for comparing square roots}
   \label{fig:result1}
@@
   \includegraphics[width=\linewidth]{l2.jpg}
-  \caption{ (Top) 64 bit (Bottom) 32 bit results }
+  \caption{ (Top) 64 bit (Bottom) 32 bit results for L2 Norm. }
   \label{fig:result1}
@@
   \includegraphics[width=\linewidth]{results1.jpg}
-  \caption{(Left) Comparison between different methods of solving linear systems: running time in seconds for Cholesky, LDLT, and CGD (with 5, 10, 15, and 20 iterations) as a function of input dimension.
+  \caption{(Left) For N = 10000: Comparison between different methods of solving linear systems: running time in seconds for Cholesky, LDLT, and CGD (with 5, 10, 15, and 20 iterations) as a function of input dimension.
 (Middle) Accuracy of CGD as a function of the input on the condition number of the input matrix $A$ with $d = 20$.
@@
   \includegraphics[width=\linewidth]{results2.jpg}
-  \caption{(Left) Comparison between different methods of solving linear systems: running time in seconds for Cholesky, LDLT, and CGD (with 5, 10, 15, and 20 iterations) as a function of input dimension.
+  \caption{(Left) For N = 1000: Comparison between different methods of solving linear systems: running time in seconds for Cholesky, LDLT, and CGD (with 5, 10, 15, and 20 iterations) as a function of input dimension.
 (Middle) Accuracy of CGD as a function of the input on the condition number of the input matrix $A$ with $d = 20$.
@@
 
-\vc{talk about outliers in these cases, skews results}
+\vc{need to replace with correct plot, found where the error came from}
 
@@
 
+3 party results on one machine
+
+\begin{figure}[h]
+  \includegraphics[width=\linewidth]{3players.jpg}
+  \caption{Results for 3 players.
+Since this is not the average of 30 runs, the lines aren't as straight yet.
+In general the results are pretty comparable to the 2 players.
+There is a noticeable overhead that each additional party incurs compared to running the same computation with less parties.
+I will include a table later to estimate the increase for each party.}
+  \label{fig:result4}
+\end{figure}
+
+4 party results on one machine
+
+\begin{figure}[h]
+  \includegraphics[width=\linewidth]{4players.jpg}
+  \caption{(Results for 4 players. Since this is not the average of 30 runs, the lines aren't as straight yet. In general the results are pretty comparable to the 2 players. }
+  \label{fig:result4}
+\end{figure}
+
 \vc{insert table comparison here??}

### 2018-03-13 paper/main.tex
Add files via upload

--- 2017-06-14 paper/main.tex
@@
 \usepackage{mathrsfs}
-
+\usepackage{float}
 \usepackage{tikz}
@@
 \author{Valerie Chen \and Valerio Pastro \and Mariana Raykova}
+\date{}
 
@@
 
-\titlepage
 \maketitle
+\section{Experiments}
 
-\pagebreak
+Our goal was to conduct a thorough investigation of the SPDZ framework in comparison to other similar frameworks of similar nature [1,2].
+The main algorithms that we considered were, LDLT Decomposition, Cholesky, Conjugate Gradient Descent (CGD), and Stochastic Gradient Descent (SGD).
+The algorithms for the first three are presented in [1] and the third is in [2].
+The experiments were run in the SPDZ framework.
+The data that we ran our experiments on were a combination of synthetic and real data sets.
+For each data set, the features were split evenly between two parties.
+The main points of consideration when running the algorithms was the precision of the results as well as the total runtime in the secure setting.
 
-\section{Preliminaries}
-\vp{stuff}
+\subsection{Cholesky, LDLT, and CGD}
 
-\section{Experiments}
-The experiments were run on the SPDZ framework, using a 128 bit and 256 bit prime field respectively.
+We primarily used a 128 bit and 256 bit prime field in SDPZ.
 In the 128 bit prime field, the bits were allotted with 28 bits after the decimal place to match the 32 bit example in the original paper and 60 bits after the decimal for the original 64 bit case.
-Compared original paper, we needed more space to the left of the decimal for comparison purposes and operations using the fixed point type.
-However, we believe this is still a fair comparison because we are not working with more precision after the decimal place and if anything using more bits should actually slow down our calculations.
-For each generated matrix of varying dimensions, the features were split evenly between two parties.
-We then wanted to see if linear decomposition algorithms like LDLT, Cholesky, and CGD (with 5, 10, 15, and 20 iterations) were able to get comparable if not better results.
+Compared to [1] we needed more space to the left of the decimal for to avoid overflow in operations using the fixed point type.
+However, this is still a fair comparison because we are not utilizing with more precision after the decimal place, and rather would actually increase run time.\\
 
-First, we ran the two algorithms two gauge time and error rates for matrices of size 10, 20, 50, 100, 200, and 500, with 200000 features each.
-Secondly, we ran on a size 20 matrices with condition numbers varying from 1 to 10.
-Both parts were completed for both bit sizes to evaluate performance differences.
-To model after the original experiment, we ran iterations of 5, 10, 15, and 20 for the CGD algorithm but added on as an extension of 25 iterations to see if necessary (i.e.
-the errors were not comparable for all sizes or condition numbers), there was still room for improvement.
-
-\section{Result}
+\noindent
+First, we ran the two algorithms two gauge time and error rates for synthetic matrices of size 10, 20, 50, 100, 200, and 500, with 10000 entries.
+As expected, utilizing more bits gives much better accuracy.
+In both cases, the error results are comparable to the garbled circuits paper.
+However, we found the original experiment truncated the generated data at 12 decimal places, however we modified it so that we could work with the full representation of the number.
+These are the average results over 30 samples with an average condition number of 1.77.
 
 \begin{figure}[h]
-  \includegraphics[width=\linewidth]{results.jpg}
-  \caption{(Left) Comparison between different methods of solving linear systems: running time in seconds for Cholesky, LDLT, and CGD (with 5, 10, 15, and 20 iterations) as a function of input dimension.
-(Middle) Accuracy of CGD as a function of the input on the condition number of the input matrix $A$ with $d = 20$.
+\centering
+  \includegraphics[scale=0.47]{results1.jpg}
+  \caption{(Left) Run time comparison in seconds as a function of input dimension.
+(Right) Accuracy as a function of the input dimensionality $d$.
+(Top) Fixed-point numbers with the $b = 80$ bits, with 60 in the fractional part.
+(Bottom) $b = 41$, with 28 in the fractional part}
+  \label{fig:result2}
+\end{figure}
+
+\noindent
+We also varied the size of $N$ to see the effect of how different numbers of entries changed the accuracy of the algorithm.
+We noticed that the condition number of the matrix varies inversely with the amount of data in the system.
+More data means smaller condition number and thus better accuracy.
+
+\begin{figure}[h]
+\centering
+  \includegraphics[scale=0.5]{results2.jpg}
+  \caption{(Left) For N = 1000: Comparison between different methods of solving linear systems: running time in seconds for Cholesky, LDLT, and CGD as a function of input dimension.
 (Right) Accuracy of Cholesky, LDLT, and CGD, as a function of the input dimensionality $d$.
@@
 (Bottom) $b = 41$, with 28 in the fractional part}
+  \label{fig:result2}
+\end{figure}
+
+\noindent
+Secondly, we ran on a size 20 and size 100 matrices with varying condition numbers from 1 to 10.
+To replicate [1] we ran iterations of 5, 10, 15, and 20 for the CGD algorithm but also tried 25 iterations to see if there were improvements.
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.6]{conditionnumber.jpg}
+  \caption{(Left) The plot of solving data with condition numbers ranging from 1 to 10 against the error, using matrices of size d = 100.
+(Right) The plot of solving data with condition numbers ranging from 1 to 10 against the error, using matrices of size d = 20.
+(Top) was ran in 32 bit SPDZ and (Bottom) was ran in 64 bit SPDZ.}
+  \label{fig:result3}
+\end{figure}
+
+\noindent
+In the condition number experiment, for CGD iterations of 5, 10, and 15, error increases as condition number increases.
+For example, in CGD 10 a small condition number may have an error of $\mathtt{10e-9}$ yet a larger condition number could have an error as large as $\mathtt{10e-3}$.
+However, in the CGD 20 case, the condition number does not have much effect due to the small errors, with the error varying only form $\mathtt{10e-11}$ to $\mathtt{10e-14}$.
+For the 60 bit precision case, we ran more iterations because there is a noticeable upward slope in the scatter plot in the CGD 20 line, indicating that still as condition numbers increased, the errors were increasing as well.
+Indeed, 25 iterations showed some improvement as we can see a flatter trend.\\
+
+\noindent
+To fine tune the CGD and Cholesky algorithms to SPDZ, we experimented with different ways of approximating square roots, to see if one or the other had better runtime or smaller error.
+The Babylonian method involves iteratively dividing and averaging an initial guess so on the next iteration can be represented as $x_{n+1} = \frac{1}{2} (x_{n} + \frac{S}{x_{n}})$.
+The Newton Method is similar but finds the square root using multiplications and not divisions: $\sqrt{S} = S \cdot (1/\sqrt{S})$.
+We originally hypothesized that using more multiplications over divisions should result in faster results.
+The results presented here are run on the Cholesky algorithm.
+However, since multiplication and division are implemented similarly in SPDZ, we did not find these results to be significant.
+
+\begin{figure}[h]
+     \centering
+  \includegraphics[scale=0.45]{sqrt.jpg}
+  \caption{ 64 bit (top) and 32 bit (bottom) results for comparing approximating square roots methods.}
   \label{fig:result1}
@@
 
-As expected, utilizing more bits gives much better accuracy.
-In both cases, the error results are comparable to the garbled circuits paper.
-However, we did find that the original experiment truncated the generated data at 12 decimal places, however we modified it so that we could work with the full representation of the number to get a better understanding of how accurate our algorithm is.
+\noindent
+For CGD, we decide to test if utilizing L2 norm vs Linfinity would make a difference in the algorithm.
+One utilizes comparisons which could be costly for arithmetic circuits and the other addition and square root.
 
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.45]{l2.jpg}
+  \caption{ (Top) 64 bit (Bottom) 32 bit results for L2 Norm. }
+  \label{fig:result1}
+\end{figure}
+
+\subsection{Experiments on multiple parties and machines}
+
+We also ran results for 3 and 4 players, which is also something that [1] considered.
+In a typical real world setting, there are often more than 2 parties, so we wanted to gauge what the additional run time would be for SPDZ to handle more parties.
+The error results were the same as those presented prior so we chose not to include them again.
+
+\begin{figure}[h]
+\centering
+  \includegraphics[scale=0.6]{multi.jpg}
+  \caption{Run time results for 3 and 4 players in 32 bits. (Left) 3 Players. (Right) 4 Players.}
+  \label{fig:result4}
+\end{figure}
+
+\noindent
+Additionally, we deployed two Amazon EC2 instances of the c4.8xlarge type and performed the same experiments as before to gauge the latency of having two players on separate machines.
+The errors were the same as before, but we observed a drastic increase in run time over a network, compared to the previous two player results.
+
+\begin{figure}[h]
+\centering
+  \includegraphics[scale=0.7]{ec2.jpg}
+  \caption{Run time results for 2 players deployed on EC2 instances averaged over 15 runs. (Left) 32 bits. (Right) 64 bits.}
+  \label{fig:result4}
+\end{figure}
+
+\subsection{Stochastic Gradient Descent}
+
+To implement SGD, we utilized the update function presented in [2]: $ w = w - \frac{1}{|B|} \alpha X^{T}_{B} \times (X_{B} \times w - Y_{B})$.
+The first dataset that we worked with was MNIST.
+However, per [2], we only trained and tested on the 0s and 1s in the dataset.
+From Figure 8, we can observe that 13 bits are sufficient enough to achieve almost perfect accuracy for prediction.
+
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.6]{mnist.jpg}
+  \caption{Varying number of bits and recreating results in [2].}
+  \label{fig:result4}
+\end{figure}
+
+\vc{Run these same results on Arcene dataset too}
+
+\noindent
+\\Using the same SGD algorithm, we also compared the resulting root mean squared error (RMSE) in 7 open sourced datasets of ranging sizes.
+
+\begin{table}[H]
+\centering
+\label{my-label}
+\begin{tabular}{@{}llllll@{}}
+\toprule
+Dataset Name          & \multicolumn{1}{c}{SGD Plaintext} & \multicolumn{2}{c}{\begin{tabular}[c]{@{}c@{}}SPDZ (28 bit)\\ error / time\end{tabular}} & \multicolumn{2}{c}{\begin{tabular}[c]{@{}c@{}}SPDZ (13 bit)\\ error / time\end{tabular}} \\ \midrule
+Student Performance   & 0.11                              & 0.12 (+8.34\%)                                  & 2.174                                  & 0.12(+8.34\%)                                   & 2.616                                  \\
+Auto MPG              & 0.56                              & 0.68 (+21.4\%)                                  & 0.663                                  & 0.68(+21.4\%)                                   & 0.704                                  \\
+Communities and Crime & 0.06                              & 0.17 (+183\%)                                   & 10.352                                 & 0.19 (+216\%)                                   & 12.363                                 \\
+Wine Quality          & 0.18                              & 0.19 (+5.55\%)                                  & 0.809                                  & 0.19 (+5.55\%)                                  & 0.969                                  \\
+Bike Sharing          & 0.23                              & 0.24 (+4.34\%)                                  & 0.910                                  & 0.25 (+8.69\%)                                  & 1.097                                  \\
+Year Prediction MSD   & 0.06                              & 0.06 (+0.0\%)                                   & 7.146                                  & 0.06 (+0.0\%)                                   & 8.513                                  \\
+Gas Sensor Array      & 0.20                              & 0.36 (+20.0\%)                                  & 1.151                                  & 0.36 (+20.0\%)                                  & 1.559                                  \\ \bottomrule
+\end{tabular}\\
+\caption{SGD Results for 7 open sourced UCI datasets.}
+\end{table}
+
+\noindent
+Running the results in a secure setting increased the RMSE in quite a few of the datasets, however it is important to note that even the increase is still quite low for large data sets.
+In comparison to CGD and Cholesky results from [1], SGD had lower RMSE even when run in SPDZ.
+
+\subsection{Analysis}
+
+\begin{table}[H]
+\centering
+\label{my-label}
+\begin{tabular}{@{}llll@{}}
+\toprule
+iterations & one machine      & two EC2      & {[}1{]} results    \\ \midrule
+5     & 206.08 & 784.74  & 10,474.57 \\
+10    & 313.07 & 1505.08 & 20,888.35 \\
+15    & 418.69 & 2039.39 & 31,300.94 \\
+20    & 527.28 & 2701.99 & n/a        \\ \bottomrule
+\end{tabular}
+\caption{Runtime comparison of SPDZ for CGD on one and two machines against [1] in seconds}
+\end{table}
+
+\noindent
 The run times for SPDZ are always comparable if not much faster.
@@
 For dimensions 100 and smaller, when Cholesky and LDLT are faster, they comparative differences are not that big (LDLT with 1.6 seconds compared to CGD 20 with 60.89 seconds) when compared to the largest differences between the larger sizes (LDLT with 1747 seconds and CGD5 with 97 seconds).
+It is also important to note that the time differences between 32 and 64 bits are not that large- only blank percent more compared to the garbled circuits result where doubling the number of bits roughly increased time by an entire magnitude of 10.
+\\
 
-\vc{insert table comparison here??}
+\noindent
+The major takeaways are that increasing the size of the matrix and increasing the condition number results in higher errors.
+To get better results, we need to run more iterations of CGD.
+For high condition numbers in large matrices, Cholesky and LDLT outperform the other methods.
+However, CGD does well in low condition numbers.
 
-In the condition number experiment, for CGD iterations of 5, 10, and 15, error increases as condition number increases.
-For example, in CGD 10 a small condition number may have an error of $\mathtt{10e-9}$ yet a larger condition number could have an error as large as $\mathtt{10e-3}$.
-However, in the CGD 20 case, the condition number does not seem to have as much of effect as the range of errors is much less., with the error varying only form $\mathtt{10e-11}$ to $\mathtt{10e-14}$ keeping in mind these are much smaller increments.
-\vc{why??
-Add some explanation about how condition number affects errors in a matrix} For the 60 bit precision case, we chose to run more iterations because in the CGD 20 line, there is a noticeable upward slope in the scatter plot, indicating that still as condition numbers increased, the errors were increasing as well.
-Indeed, 25 iterations showed some improvement as we can see a flatter trend in the points.
+\section{References}
+
+\vc{will fix this later with bibtex}
+
+[1] Privacy-Preserving Distributed Linear Regression on High-Dimensional Data
+
+[2] SecureML
 

### 2018-04-08 paper/main.tex
newest results

@@
 \centering
-  \includegraphics[scale=0.6]{mnist.jpg}
-  \caption{Varying number of bits and recreating results in [2].}
+  \includegraphics[scale=0.6]{mnistarcene.png}
+  \caption{Varying number of bits and recreating results for MNIST (left) and Arcene (right) in [2].}
   \label{fig:result4}
@@
 
-\vc{Run these same results on Arcene dataset too}
+\noindent
+The dimensions of the vector that we were predicting for MNIST was only 784, which comes from the 28 by 28 image.
+This is small compared to the size 10,000 vector that Arcene has.
+This could be an explanation for why the SGD prediction capabilities are not as good for Arcene when compared to MNIST and could explain why we need more bits in SPDZ to achieve plaintext comparable accuracy.
 

### 2018-09-18 paper/main.tex
Add files via upload

@@
 \usepackage{amsthm}
+\usepackage{amsmath}
 
@@
 
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+
+\makeatletter
+\def\BState{\State\hskip-\ALG@thistlm}
+\makeatother
+
 \input{symbols}
@@
 \maketitle
+
+\section{Background}
+
+Regression analysis and other machine learning techniques aim to build a model that fits a set of predictors to the dependent variable.
+These are commonly utilized technique in real world industries ranging from finance to medicine to security and can be crucial in understanding large scale consumer behavior.
+With the advent of big data and recent technological advances, it is becoming increasingly simple for entities to store vast amounts of user data.
+For example, individual hospitals might have their own patient data, but sharing this data for prediction or model-building purposes would be against modern day privacy laws.
+Secure multi-party computation (MPC) aims to solve this problem by providing a mechanism through which different parties and supply their data for joint computation without revealing individual values in each database.\\
+
+\noindent
+Two common regression techniques are linear and logistic regression.
+These are the fundamental building blocks for many more complex machine learning algorithms.
+In this project, our goal was to evaluate the efficacy of SPDZ as a viable secure multi-party framework for complex regression and machine learning algorithms.
+This paper also briefly explores the efficacy of neural networks in a secure computation setting.\\
+
+\noindent
+Insert stuff about SPDZ background / why we chose to use it over other things. \\
+
+\noindent
+Two previous works that this paper specifically builds off of are [1,2].
+
+\section{Algorithms}
+
+Our goal was to conduct a thorough investigation of the SPDZ framework in comparison to other similar frameworks of similar nature [1,2].
+The main algorithms that we considered were, LDLT Decomposition, Cholesky, Conjugate Gradient Descent (CGD), Stochastic Gradient Descent (SGD), Logistic Regression, and Neural Networks.
+The algorithms for the first three were first presented in [1] and the latter two in [2], but in the sections below we further detail the algorithms and modifications that were made to them to accomodate the SPDZ framework.
+The algorithms were implemented in the SPDZ framework as a proof-of-concept as well as in python as a plaintext verification of the accuracy of the secure version.
+Even though the files that is inputted into SPDZ are almost derived from python, there were capabilities that were missing in this framework that directly impacted the implementation of the algorithm.
+The following subsections detail the five algorithms that were investigated in this work as well as SPDZ specific changes that were made to make the implementation work.
+\\
+
+\noindent
+A general problem that differed SPDZ from a numpy framework in python was the lack of  vectorization, meaning there was no support for matrix operations.
+This made sense because even basic arithmetic and multiplication was difficult to implement in a secure setting.
+However, we see this as an important avenue of development due to the increasing importance of ...
+finish this sentence.
+LOOK INTO HOW MATRICES ARE WRITTEN - are there operations?
+
+\subsection{Cholesky}
+
+The Cholesky decomposition algorithm relies on having a Hermitian positive-definite matrix A and a vector b as input and outputs a factorization of the form $LL^T$ where $L$ is a lower triangular matrix with real and positive diagonal entries and $L^T$ is the conjugate transport of L, which in this case is just $L^T$.
+Then using $L$, we compute $y$ where $L^Ty = b$.
+Finally the solution $\beta$ where $A\beta = y$ is computed.
+When given a matrix that fits the Hermitian, positive-definite pattern, this method outperforms traditional LU decomposition, for solving linear equations.
+A more detailed algorithm is presented below:
+
+\begin{algorithm}[H]
+\caption{}\label{euclid}
+\begin{algorithmic}[1]
+\State $\textit{Let a be the input matrix, d be dimensions}$
+\State 1. Decompose into $LD$
+\For {j = 0, $j < d$}
+\For {k = 0, $k < j$}
+\For {i = j, $i < d$}
+\State $a[i][j] = a[i][j] - (a[i][k] *a[j][k])$.
+\EndFor
+\State $a[j][j] = sqrt(a[j][j])$
+\For {k = j+1, k < d}
+\State $a[k][j] = a[k][j] / a[j][j]$
+\EndFor 
+\EndFor
+\EndFor
+\State 2. Compute $y$ where $L^Ty = b$
+\For {i = 0, $i < d$}
+\For {j = 0, $j < i$}
+\State $b[i] = b[i] - (a[i][j] * y[i])$
+\EndFor 
+\State $y[i] = b[i] / a[i][i]$
+\EndFor
+\State 3. Compute the solution $\beta$ where $A\beta = y$
+\For {$i = d-1$, $i \geq 0$, $i = i - 1$}
+\For {j = 0, $j < i$}
+\State $b[i] = b[i] - (a[i][j] * y[i])$
+\EndFor 
+\State $y[i] = b[i] / a[i][i]$
+\EndFor
+\end{algorithmic}
+\end{algorithm}
+Runtime
+
+\subsection{LDLT}
+
+The $LDL^T$, also known as the LDL decomposition, algorithm is a close variant of the Cholesky algorithm presented in the section above.
+As the name suggests, the input matrix A is factored into $LDL^T$, where $L$ is a lower triangular matrix and $D$ is a diagonal matrix.
+We can then use $L$ and $D$ to solve for the solution to the linear equations by computing $b'$ where $Lb' = b$ and then the solution $b''$ where $Db'' = b'$.
+A detailed algorithm is detailed below:
+
+\begin{algorithm}[H]
+\caption{}\label{euclid}
+\begin{algorithmic}[1]
+\State $\textit{Let a be the input matrix, d be dimensions, b be the output vector}$
+\State 1. Decompose into $LD$
+\For {j = 0, $j < d$}
+\For {k = 0, $k < j$}
+\State $temp = a[j][k] * a[k][k]$
+\For {i = j, $i < d$}
+\State $a[i][j] = a[i][j] - (a[i][k] *temp)$.
+\EndFor
+\For {k = j+1, k < d}
+\State $a[k][j] = a[k][j] / a[j][j]$
+\EndFor 
+\EndFor
+\EndFor
+\State 2. Compute $b'$ where $Lb'= b$
+\For {i = 0, $i < d$}
+\For {j = 0, $j < i$}
+\State $b[i] = b[i] - (a[i][j] * b[j])$
+\EndFor 
+\EndFor
+\State 3. Compute $b''$ where $Db'' = b'$
+\For {i = 0, i < d}
+\State $b[i] = b[i] / a[i][i]$
+\EndFor 
+\State 4. Compute $\beta$ where $L\beta = b''$ 
+\For {$i = d-1$, $i \geq 0$, $i = i - 1$}
+\For {$j = d-1$, $j > i$, $j = j -1$}
+\State $b[i] = b[i] - (a[j][i] * b[j])$
+\EndFor 
+\EndFor
+\end{algorithmic}
+\end{algorithm}
+
+\noindent
+Runtime. The LDL variant, if efficiently implemented, requires the same space and computational complexity to construct and use but avoids extracting square roots.
+
+\subsection{Conjugate Gradient Descent}
+
+Conjugate Gradient Descent, also known as CGD, is a popular numerical method for solving linear equations of the symmetric, positive-definite form.
+The difference between CGD and the previous two algorithms is that this is iterative, which makes it particularly suitable for large systems where a direct decomposition method would be too slow.
+
+\begin{algorithm}[H]
+\caption{}\label{euclid}
+\begin{algorithmic}[1]
+\State $\textit{Let a be the input matrix, d be dimensions}$
+\State $\theta = [0]*n$
+\State $gtilda = [0]*n$
+\State $p = [0]*n$
+for
+
+\For {$i = 0$, $i < iterations$}
+\EndFor
+\end{algorithmic}
+\end{algorithm}
+
+Look at paper for algorithm modifications.
+
+Runtime why iterative makes it better
+
+\subsection{Stochastic Gradient Descent}
+
+Stochastic Gradient Descent, or SGD, \\
+
+\noindent
+Talk about batch normalization, which is also applicable to logistic regression.
+
+\begin{algorithm}[H]
+\caption{}\label{euclid}
+\begin{algorithmic}[1]
+\Procedure{MyProcedure}{}
+\State $\textit{stringlen} \gets \text{length of }\textit{string}$
+\State $i \gets \textit{patlen}$
+\BState \emph{top}:
+\If {$i > \textit{stringlen}$} \Return false
+\EndIf
+\State $j \gets \textit{patlen}$
+\BState \emph{loop}:
+\If {$\textit{string}(i) = \textit{path}(j)$}
+\State $j \gets j-1$.
+\State $i \gets i-1$.
+\State \textbf{goto} \emph{loop}.
+\State \textbf{close};
+\EndIf
+\State $i \gets i+\max(\textit{delta}_1(\textit{string}(i)),\textit{delta}_2(j))$.
+\State \textbf{goto} \emph{top}.
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Logistic Regression}
+
+Another popular regression method for two-class classification problems is logistic regression.
+The update function is a modification of the SGD update, $ w = w - \frac{1}{|B|} \alpha X^{T}_{B} \times (f(X_{B} \times w) - Y_{B})$ where $f(u)$ is the activation function $\frac{1}{1+e^{-u}}$.
+However, the exponential function is difficult to achieve in a secure setting, so one alternative is to approximate an exponential using a Taylor series and another is to use the secure computation friendly activation piecewise function defined by [2]:
+
+\[
+  f(u) =
+  \begin{cases}
+                                   0 & \text{if $u < -0.5$} \\
+                                   u + 0.5 & \text{if $-0.5 \leq u \leq 0.5$} \\
+1 & \text{if $u > 0.5$}
+  \end{cases}
+\]
+
+What distinguishes logistic regression from the previous linear regression algorithms is the use of a nonlinear activation function.
+
+\begin{algorithm}[H]
+\caption{}\label{euclid}
+\begin{algorithmic}[1]
+\Procedure{MyProcedure}{}
+\State $\textit{stringlen} \gets \text{length of }\textit{string}$
+\State $i \gets \textit{patlen}$
+\BState \emph{top}:
+\If {$i > \textit{stringlen}$} \Return false
+\EndIf
+\State $j \gets \textit{patlen}$
+\BState \emph{loop}:
+\If {$\textit{string}(i) = \textit{path}(j)$}
+\State $j \gets j-1$.
+\State $i \gets i-1$.
+\State \textbf{goto} \emph{loop}.
+\State \textbf{close};
+\EndIf
+\State $i \gets i+\max(\textit{delta}_1(\textit{string}(i)),\textit{delta}_2(j))$.
+\State \textbf{goto} \emph{top}.
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+Another challenge to implementing logistic regression is that SPDZ does not support branching using if statements, since comparing a secret value without revealing the internal value will always return, no matter the comparison on the actual value.
+To work around this, we need to convert the piecewise function into circuits.
+
+\subsection{Neural Networks}
+
+Neural networks are extensions of the more basic regression algorithms presented above, which includes multiple layers of neurons and forward and backpropagation through the layers.
+Per the experiments in [2], we used two hidden layers and 128 neurons per layer.
+
+After each layer is a nonlinear activation function and an area worth experimenting in terms of MPC is the best type of activation function, which includes
+
+We experiment with both the RELU and the square function as the activation function in the hidden layers and our proposed alternative to softmax function in the output layer.
+
+Write out the formulas... as well as detail the implementation of approximating the activation function as well as softmax.
+
+Since [2] did not detail the protocol, we provide a more in-depth description for replicating our results.
+
+\begin{algorithm}[H]
+\caption{}\label{euclid}
+\begin{algorithmic}[1]
+\Procedure{MyProcedure}{}
+\State $\textit{stringlen} \gets \text{length of }\textit{string}$
+\State $i \gets \textit{patlen}$
+\BState \emph{top}:
+\If {$i > \textit{stringlen}$} \Return false
+\EndIf
+\State $j \gets \textit{patlen}$
+\BState \emph{loop}:
+\If {$\textit{string}(i) = \textit{path}(j)$}
+\State $j \gets j-1$.
+\State $i \gets i-1$.
+\State \textbf{goto} \emph{loop}.
+\State \textbf{close};
+\EndIf
+\State $i \gets i+\max(\textit{delta}_1(\textit{string}(i)),\textit{delta}_2(j))$.
+\State \textbf{goto} \emph{top}.
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+\section{Datasets}
+
+This paper used a combination of existing, well-known and synthetically generated data.
+Many other papers (give examples) have also explored the MNIST and Arcene datasets as the canonical dataset for a proof-of-concept model.
+The MNIST is a classification dataset which consists of 60,000 black and white images of handwritten numbers from 0 to 9 scaled to a 28x28 pixel size.
+Each pixel is used as a feature which gives 784 total features in the dataset.
+The Arcene dataset is a two-class classification problem to distinguish cancerous from normal patients through mass spectrometric data.
+The mass spectrometric data is highly dimensional, with 10,000 features.
+A clarification of [2] is that we only explore both datasets in a binary setting, which is why the accuracy is so high for MNIST in particular.\\
+
+\noindent
+[1] presented a method for synthetic data generation that allowed varying of a few parameters which we took advantage to test the linear regression algorithms in different settings.
+The algorithm automatically
+
+Talk about condition numbers here... ?
+
 \section{Experiments}
 
-Our goal was to conduct a thorough investigation of the SPDZ framework in comparison to other similar frameworks of similar nature [1,2].
-The main algorithms that we considered were, LDLT Decomposition, Cholesky, Conjugate Gradient Descent (CGD), and Stochastic Gradient Descent (SGD).
-The algorithms for the first three are presented in [1] and the third is in [2].
-The experiments were run in the SPDZ framework.
+The experiments for each algorithm were run in the SPDZ framework.
 The data that we ran our experiments on were a combination of synthetic and real data sets.
@@
 The main points of consideration when running the algorithms was the precision of the results as well as the total runtime in the secure setting.
+\\
+
+\noindent
+Typically both parties were run on one machine but we also performed tests where both parties were deployed on different Amazon EC2 instances.
+It is important to note that the distribution of players across machine only affects runtime and not the accuracy of the results.
 
@@
 To implement SGD, we utilized the update function presented in [2]: $ w = w - \frac{1}{|B|} \alpha X^{T}_{B} \times (X_{B} \times w - Y_{B})$.
-The first dataset that we worked with was MNIST.
-However, per [2], we only trained and tested on the 0s and 1s in the dataset.
+The first dataset that we worked with was MNIST, which contain 28 by 28 pixels of handwritten numbers and their corresponding labels.
+While the numbers in MNIST range from 0 to 9, [2] turned it into a binary problem by only using the handwritten 0 and 1s.
+The second dataset was Arcene which is also a two-class problem of cancer versus noncancerous spectrometric data.
 From Figure 8, we can observe that 13 bits are sufficient enough to achieve almost perfect accuracy for prediction.
@@
 Bike Sharing          & 0.23                              & 0.24 (+4.34\%)                                  & 0.910                                  & 0.25 (+8.69\%)                                  & 1.097                                  \\
+Blog Feedback   & 0.04                              & 0.04 (+0.0\%)                                   & 9.455                                  & 0.04 (+0.0\%)                                   & 9.091                                  \\
+CT Slices   & 0.22                              & 0.22 (+0.0\%)                                   & 16.458                                  & 0.22 (+0.0\%)                                   & 16.295                                  \\
 Year Prediction MSD   & 0.06                              & 0.06 (+0.0\%)                                   & 7.146                                  & 0.06 (+0.0\%)                                   & 8.513                                  \\
@@
 \end{tabular}\\
-\caption{SGD Results for 7 open sourced UCI datasets.}
+\caption{SGD Results for 9 open sourced UCI datasets.}
 \end{table}
@@
 In comparison to CGD and Cholesky results from [1], SGD had lower RMSE even when run in SPDZ.
-
-\subsection{Analysis}
 
@@
 
+\subsection{Logistic Regression}
+
+\noindent
+With logistic regression, the datasets that we looked at were MNIST and Arcene.
+In their experiments, they found that the new function outperformed a polynomial approximation of degree 10.
+We verify these results in SPDZ.
+The idea behind why this piecewise functions outperforms a polynomial approximation is in the tail behavior.
+From the results below, we observe that the new activation function is more stable than a polynomial approximation.
+For different
+
+\begin{table}[H]
+\centering
+\label{my-label}
+\begin{tabular}{@{}lllllll@{}}
+\toprule
+       & Plaintext & New Activation function & \multicolumn{4}{l}{Polynomial Approximation} \\
+       &           &                         & degree 2  & degree 5  & degree 7 & degree 10 \\ \midrule
+MNIST  & 99.9\%    & 95\%                    & 97\%      & 85\%      & 91\%     & 99.5\%    \\
+Arcene & 72.0\%    &           44.0\%              &   44.0\%        &    44.5\%       &    65\%      &    72\%       \\ \bottomrule
+\end{tabular}
+\caption{Comparing the validation accuracy for different activation functions for logistic regression for a fixed train/test dataset using 32 bits (for 1 epoch and learning rate of 0.01)}
+\end{table}
+
+I'd like to note that in plaintext, using the new activation function, the accuracy is only 56\%.
+MNIST contains thousands of entries and from the SGD input.
+For Arcene, running more iterations would improve accuracy.
+For MNIST it would not.
+MNIST is 200 by 784 whereas Arcene is 200 by 10,000.
+This result was run for 20 bits of precision after the decimal.
+For MNIST, you don't have to train on entire dataset...
+(and SPDZ will overflow/seg fault if try to feed the thousands in..)
+
+\begin{table}[H]
+\centering
+\label{my-label}
+\begin{tabular}{@{}lcccc@{}}
+\toprule
+                        & \multicolumn{2}{c}{MNIST} & \multicolumn{2}{c}{Arcene} \\ \midrule
+                        & Compile      & Run        & Compile      & Run         \\
+New Activation Function & 26.01        & 18.8       & 220.14       & 224.47      \\
+2 Polynomial            & 27.79        & 19.28      & 222.85       & 239.98      \\
+5 Polynomial            & 30.32        & 19.8       & 225.33       & 241.59      \\
+7 Polynomial            & 31.99        & 20.65      & 228.33       & 239.11      \\
+10 Polynomial           & 34.24        & 20.71      & 230.81       & 240.79      \\ \bottomrule
+\end{tabular}
+\caption{The runtimes for the results in table 2}
+\end{table}
+
+Choosing the best results from Table 2, we selected the degree 10 polynomial approximation for both the MNIST and Arcene datasets.
+
+\vc{another idea for results: do the same thing with varying bits and iterations.. }
+
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.6]{mnistarcene.png}
+  \caption{REPLACE THIS.}
+  \label{fig:result4}
+\end{figure}
+
+\subsection{Neural Networks}
+
+\section{Conclusion and Future Work}
+
+Not sure what to write here..
+
 \section{References}

### 2018-10-02 paper/nips_format.tex
nips format

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2018

% ready for submission
%\usepackage{nips_2018}

% to compile a preprint version, e.g., for submission to arXiv, add
% add the [preprint] option:
 \usepackage[preprint]{nips_2018}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2018}

% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2018}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{multicol,color}

\title{Experimental Secure Multi-Party Computation on Real Data using SPDZ}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  Valerie Chen\\
  Department of Computer Science\\
  Yale University\\
  %% Address?? \\
  \texttt{v.chen@yale.edu} \\
  %% examples of more authors
  \And
  Mariana Raykova \\
  Department of Computer Science\\
  Yale University \\
  %% Address?? \\
  \texttt{mariana.raykova@yale.edu} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}

\maketitle

\begin{abstract}
  Secure Multi-Party Computation (MPC) is an area of theoretical cryptography research that has the power to revolutionize computation between shared parties holding private data. However theoretical protocols are often not scalable and efficient given large quantities of real world data. This project investigates an MPC protocol called SPDZ secure against malicious adversaries in comparison to two other semi-honest MPC protocols Obliv-C and SecureML in machine learning and regression applications. We demonstrate its potential to be used in real world applications with experiments to show that it outperforms the other two methods while maintaining a more rigorous level of security.
\end{abstract}

\section{Introduction}

Regression analysis and other machine learning techniques aim to build a model that fits a set of predictors to the dependent variable. Models built from a large set of collected data can be used to predict future behavior. With the advent of big data, it is becoming increasingly common for entities to store vast amounts of user data, run regression analyses and build models of this data to gain insights. A desired use case could be, for example, the Center for Disease Control is interested in identifying disease outbreak and many individual hospitals have their own patient data which could benefit this study. The problem is that openly sharing this data for prediction or model-building purposes would be against modern day privacy laws as it would leak private individual data. This application of secure MPC is one that would be extremely valuable in today\textquotesingle s world.

\subsection{Multi-Party Computation}

Secure MPC addresses this problem by providing a mechanism through which different parties supply their data for joint computation without revealing individual values from each database or even intermediate values. This need for MPC motivates our study to bridge theory and application through experiments on real-world data.

In general, MPC protocols predominantly come in two forms: security against semi-honest and malicious adversaries. In semi-honest security, a more naive model, it is assumed that the party will follow the protocol as specified but will merely gather information intermediate values. Typically semi-honest security is more efficient on computation as it has fewer requirements. In malicious security, dishonest parties may attempt to deviate from the specified protocol, and the protocol must be able to detect such cheating and abort the computation - thus any successful output is guaranteed to have privacy preserved. This notion can be rigorously proved.

For this study, we chose the SPDZ framework, which ensures against malicious security, and compare against two semi-honest methods Obliv-C and SecureML. SPDZ performs the checks necessary to detect dishonest parties through using random data in the hopes of making computation more efficient. Obliv-C is based on Yao\textquotesingle s garbled circuits and introduces optimizations on inner products to speed up regression calculation. SecureML\textquotesingle s focus is with developing MPC-friendly alternatives to non-linear activation functions which are often intensive to compute.

Since secure computation slower and could be limited in accuracy when compared to computation in plaintext, there are a number of modifications to plaintext algorithms that can be made to obtain a reasonable degree of accuracy in more efficient time.

\section{ML Functionalities}

Our goal was to conduct a thorough investigation of the SPDZ framework in the applications of regression and machine learning algorithms in a similar fashion to that done in  [1,2], evaluating in terms of runtime and accuracy. The main algorithms that we considered were, LDLT Decomposition, Cholesky, Conjugate Gradient Descent (CGD), Stochastic Gradient Descent (SGD), Logistic Regression. The algorithms were implemented in the SPDZ framework as a proof-of-concept as well as in python as a plaintext verification of the accuracy of the secure version.

\subsection{Direct vs. Iterative Decomposition}

LDLT and Cholesky are both variants of direct decomposition methods which decompose a Hermitian, positive-definite matrix into a lower triangular matrix and its conjugate transpose. The algorithms are cubic in complexity with asymptotic run time of $O(d^3)$. This can become difficult to work with when looking at real world data that has large number of features and entries. The difference between LDLT and Cholesky is that Cholesky requires a square root. We looked at two different methods to approximate the square root function in SPDZ. The Babylonian method involves iteratively dividing and averaging an initial guess so on the next iteration can be represented as $x_{n+1} = \frac{1}{2} (x_{n} + \frac{S}{x_{n}})$. The Newton Method is similar but finds the square root using multiplications and not divisions: $\sqrt{S} = S \cdot (1/\sqrt{S})$.

In terms of an iterative approach to regression, [1] proposed a modification for the normalization in the iterative version of CGD that would be more stable with fixed-point numbers, which is detailed here:

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.45]{cgd_alg}
  \caption{Iterative fixed-point algorithm for CGD}
\end{figure}
We applied the same algorithm to SPDZ to see if the same modification would work. The goal of comparing iterative versus a direct method would be to justify the potential accuracy trade off for a much smaller run time.

\subsection{Gradient Descent}

Stochastic gradient descent is also an iterative method to minimize a differentiable cost function and is the mechanism behind most neural networks. In linear regression the cost function is derived from the mean squared error. The update consists of forward propagation to calculate the predicted output $y_i^* = x_i \cdot w$ and a backward propagation phase to adjust weights accordingly given differential error and learning rate $\alpha(y_i^*-y_i)x_{ij}$ for an update function of $w_j = w_j  - \alpha(x_i \cdot w *-y_i)x_{ij}$.

We implemented mini-batch SGD so the weights would converge more quickly and in smoother fashion to the minimum. Since SPDZ had limited functionality for matrix manipulations and vectorization, we were not able to take advantage of the time speed-ups in this sense. The update function for the batched version is: 
\[
 w = w - \frac{1}{|B|} \alpha X^{T}_{B} \times (X_{B} \times w - Y_{B})
\]

\subsubsection{Non-Linear Activations}

For logistic regression, the activation function is defined as the logistic function $f(u) = \frac{1}{1+e^-u}$. Using the cross entropy function, we derive a similar update function for mini-batched SGD as $ w = w - \frac{1}{|B|} \alpha X^{T}_{B} \times f(X_{B} \times w - Y_{B})$. To approximate the activation function, [2] proposed this following piecewise function:

\[
  f(u) =
  \begin{cases}
                                   0 & \text{if $u < -0.5$} \\
                                   u + 0.5 & \text{if $-0.5 \leq u \leq 0.5$} \\
1 & \text{if $u > 0.5$}
  \end{cases}
\]

We compare the results of this MPC-friendly piecewise function to a more standard approximation approach of taking the Taylor Series expansion to varying degrees.

\section{Experiments}

\subsection{Experimental Setup}

The main metrics of evaluations were the error in accuracy and the RMSE for prediction error. We varied the precision after the decimal point (32 and 64 bits) or less depending on the comparison.

To simulate the players, we needed to identify a source of data. We used both a combination of existing, well-known datasets (MNIST, Arcene, and 9 other UCI open-source datasets) as well as synthetically generated data. The advantage to using synthetic data is being able to control characteristics of the data like the dimensions (d = 10, 20, 50, 100, 200, 500), condition number (cd = [1,10]), and number of examples (n = 1000, 100000).

Typically in a two player setting, both parties were run on one machine but we also performed tests where both parties were deployed on separate Amazon EC2 instances. It is important to note that the distribution of players across machine only affects runtime and not the accuracy of the results. We also looked at up to four parties as a form of validation of the SPDZ protocol.

\subsection{Results}

This section presents empirical results for real-world and synthetic data evaluated on the SPDZ protocol, comparing the five different algorithms in terms of accuracy and run time for various parameters.

For LDLT, Cholesky and various iterations of CGD, we evaluated on synthetically generated data of varying sizes and condition numbers. The direct decomposition methods grew exponentially in run time on the log scale as the matrix size increased, which means that it is not suitable for large size real data. Alternatively, the iterative CGD runtime increases at a much slower rate. In the middle column of Figure 2, we find that about 20 iterations is sufficient to reach maximum accuracy given the number of allocated bits.

\begin{figure}
\centering
  \includegraphics[scale=0.4]{allregression.png}
  \label{fig:result4}
    \caption{(Left) Run time comparison in seconds as a function of input dimension. (Middle) Condition number as a function of accuracy. (Right) Accuracy as a function of the input dimensionality $d$. (Top) Fixed-point numbers with the $b = 80$ bits, with 60 in the fractional part. (Bottom) $b = 41$, with 28 in the fractional part}
\end{figure}

Figure 3, which compares MNIST and Arcene results, demonstrates that the number of bits of precision needed to get good accuracy is highly dependent on the dataset. In terms of logistic regression, for SPDZ, we did not find that the new activation function was a better alternative to taking a Taylor Series approximation for the exponential function as shown in Table 1.

\begin{figure}
\centering
  \includegraphics[scale=0.6]{mnistarcene.png}
  \label{fig:result4}
   \caption{Varying number of bits and recreating results for MNIST (left) and Arcene (right) in [2].}
\end{figure}

\begin{table}
\caption{Comparing the validation accuracy for different activation functions for logistic regression for a fixed train/test dataset using 32 bits (for 1 epoch and learning rate of 0.01)}
\centering
\label{my-label}
\begin{tabular}{@{}lllllll@{}}
\toprule
       & Plaintext & New Activation function & \multicolumn{4}{l}{Polynomial Approximation} \\
       &           &                         & degree 2  & degree 5  & degree 7 & degree 10 \\ \midrule
MNIST  & 99.9\%    & 95\%                    & 97\%      & 85\%      & 91\%     & 99.5\%    \\
Arcene & 72.0\%    &           44.0\%              &   44.0\%        &    44.5\%       &    65\%      &    72\%       \\ \bottomrule
\end{tabular}
\end{table}

\subsection{Comparison}

\begin{table}
\caption{Runtime results against }
\centering
\begin{tabular}{@{}llcl@{}}
\toprule
\multicolumn{1}{c}{Datasets}                                           & \multicolumn{3}{c}{Runtime (seconds)}                                      \\ \midrule
CGD (64 bit)                                                           & \multicolumn{1}{c}{Ours} & Ours (EC2) & \multicolumn{1}{c}{Obliv-C}        \\ \midrule
d=100                                                                  & 144                      & 243        & \textgreater $10^4$ \\ \midrule
d=500                                                                  & 527                      & 2701       & \textgreater $10^3$ \\ \midrule
\multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}SGD: \\ MNIST\end{tabular}} & \multicolumn{1}{c}{Ours} & \multicolumn{2}{c}{SecureML}                    \\ \cmidrule(l){2-4} 
                                                                       & 10200                    & \multicolumn{2}{c}{10330}                       \\ \bottomrule
\end{tabular}
\end{table}

Through this work, we demonstrated that SPDZ outperforms other protocols Obliv-C and SecureML in terms of accuracy, as evidenced in Table FILL, efficiency, and security. In future work, we plan to expand the selected algorithms beyond linear and logistic regression to include neural networks and deep neural networks. This investigation into the applicability of secure MPC on real world data demonstrates its potential change the way data is used in a variety of applications in today\textquotesingle s world.

\subsubsection*{Acknowledgments}

Who to put here / how long should this be / does this count in page limit?

\section*{References}

\small

[1] Gasc\'{o}n, A., Schoppmann, P., Balle, B., Raykova, M., Doerner, J., Zahur, S., \& Evans, D. (2017). Privacy-preserving distributed linear regression on high-dimensional data. {\it Proceedings on Privacy Enhancing Technologies, 2017(4)}, pp.\ 345-364.

[2] Mohassel, P., \& Zhang, Y. (2017). SecureML: A system for scalable privacy-preserving machine learning. {\it 38th IEEE Symposium on Security and Privacy (SP)}, pp. 19-38. IEEE.

\end{document}

### 2018-10-02 paper/nips_format.tex
Add files via upload

@@
 % ready for submission
-%\usepackage{nips_2018}
+\usepackage{nips_2018}
 
@@
 % add the [preprint] option:
- \usepackage[preprint]{nips_2018}
+ %\usepackage[preprint]{nips_2018}
 
@@
   Valerie Chen\\
-  Department of Computer Science\\
   Yale University\\
@@
   \texttt{v.chen@yale.edu} \\
+  \And
+  Valerio Pastro\\
+  Yale University\\
+  \textt{valerio.pastro@yale.edu}\\
   %% examples of more authors
@@
   Mariana Raykova \\
-  Department of Computer Science\\
   Yale University \\
@@
 
-In terms of an iterative approach to regression, [1] proposed a modification for the normalization in the iterative version of CGD that would be more stable with fixed-point numbers, which is detailed here:
+In terms of an iterative approach to regression, [1] proposed a modification for the normalization in the iterative version of CGD that would be more stable with fixed-point numbers.
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.45]{cgd_alg}
-  \caption{Iterative fixed-point algorithm for CGD}
-\end{figure}
 We applied the same algorithm to SPDZ to see if the same modification would work.
@@
 
-\subsubsection*{Acknowledgments}
-
-Who to put here / how long should this be / does this count in page limit?
-
 \section*{References}

### 2018-10-05 paper/nips_format.tex
no authors
newest draft

@@
 % to compile a camera-ready version, add the [final] option, e.g.:
-% \usepackage[final]{nips_2018}
+%\usepackage[final]{nips_2018}
 
@@
 However theoretical protocols are often not scalable and efficient given large quantities of real world data.
-This project investigates an MPC protocol called SPDZ secure against malicious adversaries in comparison to two other semi-honest MPC protocols Obliv-C and SecureML in machine learning and regression applications.
+This project investigates an MPC protocol called SPDZ which is secure against malicious adversaries in comparison to two other semi-honest MPC protocols Obliv-C and SecureML in machine learning and regression applications.
 We demonstrate its potential to be used in real world applications with experiments to show that it outperforms the other two methods while maintaining a more rigorous level of security.
@@
 A desired use case could be, for example, the Center for Disease Control is interested in identifying disease outbreak and many individual hospitals have their own patient data which could benefit this study.
-The problem is that openly sharing this data for prediction or model-building purposes would be against modern day privacy laws as it would leak private individual data.
+The problem is that openly sharing this data for prediction or model-building purposes would be against modern day privacy laws as it would leak private individual data [2].
 This application of secure MPC is one that would be extremely valuable in today\textquotesingle s world.
@@
 
-In general, MPC protocols predominantly come in two forms: security against semi-honest and malicious adversaries.
+In general, MPC protocols predominantly come in two forms: security against semi-honest and malicious adversaries [4].
 In semi-honest security, a more naive model, it is assumed that the party will follow the protocol as specified but will merely gather information intermediate values.
@@
 
-For this study, we chose the SPDZ framework, which ensures against malicious security, and compare against two semi-honest methods Obliv-C and SecureML.
+For this study, we chose the SPDZ framework [1,5], which ensures against malicious security, and compare against two semi-honest methods Obliv-C [7] and SecureML [6].
 SPDZ performs the checks necessary to detect dishonest parties through using random data in the hopes of making computation more efficient.
-Obliv-C is based on Yao\textquotesingle s garbled circuits and introduces optimizations on inner products to speed up regression calculation.
-SecureML\textquotesingle s focus is with developing MPC-friendly alternatives to non-linear activation functions which are often intensive to compute.
+Obliv-C is based on Yao\textquotesingle s garbled circuits and [3] introduces optimizations on inner products and algorithmic modifications to speed up regression calculation.
+SecureML\textquotesingle s focus is with developing MPC-friendly alternatives to non-linear activation functions which are often intensive to compute as well as vectorization techniques in a shared setting.
 
@@
 
-Our goal was to conduct a thorough investigation of the SPDZ framework in the applications of regression and machine learning algorithms in a similar fashion to that done in  [1,2], evaluating in terms of runtime and accuracy.
+Our goal was to conduct a thorough investigation of the SPDZ framework in the applications of regression and machine learning algorithms in a similar fashion to that done in  [3,6], evaluating in terms of runtime and accuracy.
 The main algorithms that we considered were, LDLT Decomposition, Cholesky, Conjugate Gradient Descent (CGD), Stochastic Gradient Descent (SGD), Logistic Regression.
@@
 The algorithms are cubic in complexity with asymptotic run time of $O(d^3)$.
-This can become difficult to work with when looking at real world data that has large number of features and entries.
+This can become inefficient to work with when looking at real world data that has large number of features and entries.
 The difference between LDLT and Cholesky is that Cholesky requires a square root.
-We looked at two different methods to approximate the square root function in SPDZ.
-The Babylonian method involves iteratively dividing and averaging an initial guess so on the next iteration can be represented as $x_{n+1} = \frac{1}{2} (x_{n} + \frac{S}{x_{n}})$.
-The Newton Method is similar but finds the square root using multiplications and not divisions: $\sqrt{S} = S \cdot (1/\sqrt{S})$.
+In MPC, non-linear functions like square roots are difficult to represent.
+Thus we looked at an iterative Newton method as a way of approximating the square root in SPDZ to solve for $x_i$ where $x_{i}^2 = S$:
 
-In terms of an iterative approach to regression, [1] proposed a modification for the normalization in the iterative version of CGD that would be more stable with fixed-point numbers.
+\[
+x_{n+1} = \frac{1}{2} (x_{n} + \frac{S}{x_{n}})
+\]
 
+In terms of an iterative approach to regression, [3] proposed a modification for the normalization in the iterative version of CGD that would be more stable with fixed-point numbers.
 We applied the same algorithm to SPDZ to see if the same modification would work.
@@
 
-Stochastic gradient descent is also an iterative method to minimize a differentiable cost function and is the mechanism behind most neural networks.
+Stochastic gradient descent is also an iterative method that minimizes a differentiable cost function and is the driving mechanism behind most neural networks.
 In linear regression the cost function is derived from the mean squared error.
@@
 
+Logistic regression models a binary dependent variable.
 For logistic regression, the activation function is defined as the logistic function $f(u) = \frac{1}{1+e^-u}$.
@@
 
-Typically in a two player setting, both parties were run on one machine but we also performed tests where both parties were deployed on separate Amazon EC2 instances.
+Typically in a two player setting, both parties were run on one machine but we also performed tests where both parties were deployed on separate Amazon EC2 m4.large instances.
 It is important to note that the distribution of players across machine only affects runtime and not the accuracy of the results.
@@
 
-For LDLT, Cholesky and various iterations of CGD, we evaluated on synthetically generated data of varying sizes and condition numbers.
+For LDLT, Cholesky and various iterations of CGD, we evaluated on synthetically generated data of varying sizes and condition numbers, since the data needed to be of a specific form.
 The direct decomposition methods grew exponentially in run time on the log scale as the matrix size increased, which means that it is not suitable for large size real data.
 Alternatively, the iterative CGD runtime increases at a much slower rate.
-In the middle column of Figure 2, we find that about 20 iterations is sufficient to reach maximum accuracy given the number of allocated bits.
+In the middle column of Figure 2, we find that about 20 iterations is sufficient to reach maximum accuracy given the number of allocated bits even with varying condition numbers, meaning the data might not be as well-formed.
+Particularly for the 64-bit case, shown on the bottom right, the accuracy is identical for CGD after 15 iterations and Cholesky/LDLT.
 
-\begin{figure}
+\begin{figure}[h!]
 \centering
@@
 
-Figure 3, which compares MNIST and Arcene results, demonstrates that the number of bits of precision needed to get good accuracy is highly dependent on the dataset.
-In terms of logistic regression, for SPDZ, we did not find that the new activation function was a better alternative to taking a Taylor Series approximation for the exponential function as shown in Table 1.
+Figure 2, which compares MNIST and Arcene results, demonstrates that the number of bits of precision needed to get good accuracy is highly dependent on the dataset.
+For MNIST, 13 bits was sufficient to match plaintext accuracy but 28 bits were needed for Arcene.
+The MNIST data contains only 784 features while there are 10,000 in the Arcene data, thus the larger the feature set the more bits needed.
+While the numbers in the MNIST data ranged from 0 to 9, [6] only used the 0s and 1s labels from the dataset, thus reducing it to a binary problem.
+We also compared the root mean squared error (RMSE) of SGD on 9 UCI open-sourced datasets of ranging sizes to results in [3] and found lower RMSE results.
+Due to limited space, these results are not presented.
 
-\begin{figure}
+\begin{figure}[h!]
 \centering
@@
 
-\begin{table}
+In terms of logistic regression, for SPDZ, we did not find that the new activation function was a better alternative to taking a Taylor Series approximation for the exponential function as shown in Table 1.
+We found that for SPDZ, which is based on arithmetic circuits, the extra time to take a few extra degrees in the approximation was negligible.
+
+\begin{table}[h!]
 \caption{Comparing the validation accuracy for different activation functions for logistic regression for a fixed train/test dataset using 32 bits (for 1 epoch and learning rate of 0.01)}
@@
 
-\begin{table}
+\begin{table}[h!]
 \caption{Runtime results against }
@@
 
-Through this work, we demonstrated that SPDZ outperforms other protocols Obliv-C and SecureML in terms of accuracy, as evidenced in Table FILL, efficiency, and security.
+Through this work, we demonstrated that SPDZ outperforms other protocols Obliv-C and SecureML in terms of accuracy, as evidenced in Table 2, efficiency, and security.
 In future work, we plan to expand the selected algorithms beyond linear and logistic regression to include neural networks and deep neural networks.
@@
 
-[1] Gasc\'{o}n, A., Schoppmann, P., Balle, B., Raykova, M., Doerner, J., Zahur, S., \& Evans, D.
+[1] Damg\aa rd, I., Keller, M., Larraia, E., Pastro, V., Scholl, P., \& Smart, N.
+P.
+(2013).
+Practical covertly secure MPC for dishonest majority?or: breaking the SPDZ limits.
+{\it European Symposium on Research in Computer Security}, pp.\ 1-18.
+
+[2] Du, W., \& Atallah, M.
+J.
+(2001).
+Secure multi-party computation problems and their applications: a review and open problems.
+{\it Proceedings of the 2001 workshop on New security paradigms}, pp.\ 13-22.
+ACM.
+
+[3] Gasc\'{o}n, A., Schoppmann, P., Balle, B., Raykova, M., Doerner, J., Zahur, S., \& Evans, D.
 (2017).
@@
 
-[2] Mohassel, P., \& Zhang, Y. (2017). SecureML: A system for scalable privacy-preserving machine learning. {\it 38th IEEE Symposium on Security and Privacy (SP)}, pp. 19-38. IEEE.
+[4] Hazay, C., \& Lindell, Y. (2010). A Note on the Relation between the Definitions of Security for Semi-Honest and Malicious Adversaries. {\it IACR Cryptology ePrint Archive}, pp.\ 551.
+
+[5] Keller, M., Scholl, P., \& Smart, N.
+P.
+(2013).
+An architecture for practical actively secure MPC with dishonest majority.
+{\it Proceedings of the 2013 ACM SIGSAC conference on Computer \& communications security}, pp.\ 549-560.
+ACM.
+
+[6] Mohassel, P., \& Zhang, Y. (2017). SecureML: A system for scalable privacy-preserving machine learning. {\it 38th IEEE Symposium on Security and Privacy (SP)}, pp. 19-38. IEEE.
+
+[7] Zahur, S., \& Evans, D. (2015). Obliv-C: A Language for Extensible Data-Oblivious Computation. {\it IACR Cryptology ePrint Archive, 2015}, 1153.
 

### 2018-10-07 paper/nips_format.tex
added revisions

@@
 \begin{abstract}
-  Secure Multi-Party Computation (MPC) is an area of theoretical cryptography research that has the power to revolutionize computation between shared parties holding private data.
+  Secure Multi-Party Computation (MPC) is an area of cryptography research that has the power to revolutionize computation between shared parties holding private data.
 However theoretical protocols are often not scalable and efficient given large quantities of real world data.
@@
 
-In general, MPC protocols predominantly come in two forms: security against semi-honest and malicious adversaries [4].
+In general, MPC protocols predominantly come in two forms: security against semi-honest [6] and malicious adversaries [4].
 In semi-honest security, a more naive model, it is assumed that the party will follow the protocol as specified but will merely gather information intermediate values.
@@
 
-For this study, we chose the SPDZ framework [1,5], which ensures against malicious security, and compare against two semi-honest methods Obliv-C [7] and SecureML [6].
+For this study, we chose the SPDZ framework [1,5], which ensures against malicious security, and compare against two semi-honest methods Obliv-C [9] and SecureML [7].
 SPDZ performs the checks necessary to detect dishonest parties through using random data in the hopes of making computation more efficient.
@@
 LDLT and Cholesky are both variants of direct decomposition methods which decompose a Hermitian, positive-definite matrix into a lower triangular matrix and its conjugate transpose.
-The algorithms are cubic in complexity with asymptotic run time of $O(d^3)$.
+The algorithms are cubic in complexity with asymptotic run time of $O(d^3)$, where $d$ is the dimension of the input matrix.
 This can become inefficient to work with when looking at real world data that has large number of features and entries.
@@
 In linear regression the cost function is derived from the mean squared error.
-The update consists of forward propagation to calculate the predicted output $y_i^* = x_i \cdot w$ and a backward propagation phase to adjust weights accordingly given differential error and learning rate $\alpha(y_i^*-y_i)x_{ij}$ for an update function of $w_j = w_j  - \alpha(x_i \cdot w *-y_i)x_{ij}$.
+The update consists of forward propagation to calculate the predicted output $y_i^* = x_i \cdot w$ and a backward propagation phase to adjust weights accordingly given differential error and learning rate $\alpha(y_i^*-y_i)x_{ij}$ for an update function of $w_j = w_j  - \alpha(x_i \cdot w *-y_i)x_{ij}$, where $w$ are the weights, $\alpha$ the learning rate, $x$ the input, and $y$ the expected output [7].
 
 We implemented mini-batch SGD so the weights would converge more quickly and in smoother fashion to the minimum.
-Since SPDZ had limited functionality for matrix manipulations and vectorization, we were not able to take advantage of the time speed-ups in this sense.
-The update function for the batched version is:
+Instead of selecting one sample per iteration, a small batch of size $B$ samples are selected and the update function is performed collectively.
+Since SPDZ had limited functionality for matrix manipulations and vectorization, we were not able to take advantage of the time speed-ups in this sense, but were still able to take advantage of the accuracy benefit.
+The update function for the batched version, where $w$ are the weights, $\alpha$ the learning rate, $X_{B}$ the batched input, and $Y_{B}$ the corresponding expected batched output:
 \[
@@
 
-The main metrics of evaluations were the error in accuracy and the RMSE for prediction error. We varied the precision after the decimal point (32 and 64 bits) or less depending on the comparison.
+The main metrics of evaluations were the latency in evaluation in a privacy-preserving setting and the error in accuracy, to see what trade-offs and modifications should be made to achieve good accuracy while optimizing efficiency.
+We varied the precision after the decimal point (32 and 64 bits) or less depending on the comparison.
 
-To simulate the players, we needed to identify a source of data.
+To evaluate the SPDZ protocol implementation, we needed to identify a source of data.
 We used both a combination of existing, well-known datasets (MNIST, Arcene, and 9 other UCI open-source datasets) as well as synthetically generated data.
-The advantage to using synthetic data is being able to control characteristics of the data like the dimensions (d = 10, 20, 50, 100, 200, 500), condition number (cd = [1,10]), and number of examples (n = 1000, 100000).
+The advantage to using synthetic data is being able to control characteristics of the data like the dimensions ($d$ = 10, 20, 50, 100, 200, 500), condition number ($cd$ = [1,10]), and number of examples ($n$ = 1000, 100000).
 
-Typically in a two player setting, both parties were run on one machine but we also performed tests where both parties were deployed on separate Amazon EC2 m4.large instances.
+In our experiments, both parties were mostly run on one machine to simplify the procedure, but we also performed tests where both parties were deployed on separate Amazon EC2 m4.large instances.
 It is important to note that the distribution of players across machine only affects runtime and not the accuracy of the results.
@@
 
-For LDLT, Cholesky and various iterations of CGD, we evaluated on synthetically generated data of varying sizes and condition numbers, since the data needed to be of a specific form.
+\begin{figure}[h!]
+\centering
+  \includegraphics[scale=0.4]{allregression.png}
+  \label{fig:result4}
+    \caption{(Left) Run time as a function of input dimension.
+(Middle) Condition number as a function of accuracy.
+(Right) Accuracy as a function of the input dimension.
+(Top) Fixed-point with 60 bits of precision.
+(Bottom) Fixed-point with 28 bits of precision}
+\end{figure}
+
+For LDLT, Cholesky and various iterations of CGD, we evaluated on synthetically generated data of varying sizes and condition numbers.
+The larger the condition number, the larger the error in approximations of the solution.
 The direct decomposition methods grew exponentially in run time on the log scale as the matrix size increased, which means that it is not suitable for large size real data.
@@
 \begin{figure}[h!]
+\vspace{-4mm}
 \centering
-  \includegraphics[scale=0.4]{allregression.png}
+  \includegraphics[scale=0.6]{mnistarcene.png}
   \label{fig:result4}
-    \caption{(Left) Run time comparison in seconds as a function of input dimension.
-(Middle) Condition number as a function of accuracy.
-(Right) Accuracy as a function of the input dimensionality $d$.
-(Top) Fixed-point numbers with the $b = 80$ bits, with 60 in the fractional part.
-(Bottom) $b = 41$, with 28 in the fractional part}
+  \vspace{-4mm}
+   \caption{Comparing accuracy of privacy preserving linear regression with fixed point and plaintext training on floating point for MNIST (left) and Arcene (right) in [2].}
+   \vspace{-4mm}
 \end{figure}
@@
 For MNIST, 13 bits was sufficient to match plaintext accuracy but 28 bits were needed for Arcene.
-The MNIST data contains only 784 features while there are 10,000 in the Arcene data, thus the larger the feature set the more bits needed.
-While the numbers in the MNIST data ranged from 0 to 9, [6] only used the 0s and 1s labels from the dataset, thus reducing it to a binary problem.
+The MNIST data contains only 784 features while there are 10,000 in the Arcene data, 3,000 of which are considered "probes" with no predictive power, which could explain the lower overall accuracy [8].
+While the numbers in the MNIST data ranged from 0 to 9, to match the approach in [7], we only used the 0s and 1s labels from the dataset, thus reducing it to a binary problem.
+We did run the computation with all 10 digits, but found that gradient descent achieved a much lower accuracy of about 19\%.
 We also compared the root mean squared error (RMSE) of SGD on 9 UCI open-sourced datasets of ranging sizes to results in [3] and found lower RMSE results.
 Due to limited space, these results are not presented.
-
-\begin{figure}[h!]
-\centering
-  \includegraphics[scale=0.6]{mnistarcene.png}
-  \label{fig:result4}
-   \caption{Varying number of bits and recreating results for MNIST (left) and Arcene (right) in [2].}
-\end{figure}
 
@@
 \begin{table}[h!] 
-\caption{Comparing the validation accuracy for different activation functions for logistic regression for a fixed train/test dataset using 32 bits (for 1 epoch and learning rate of 0.01)}
+\caption{Comparing the validation accuracy for different activation functions for logistic regression.}
 \centering
@@
 \end{tabular}
+\vspace{-4mm}
 \end{table}
@@
 
-\begin{table}[h!]
-\caption{Runtime results against }
-\centering
-\begin{tabular}{@{}llcl@{}}
-\toprule
-\multicolumn{1}{c}{Datasets}                                           & \multicolumn{3}{c}{Runtime (seconds)}                                      \\ \midrule
-CGD (64 bit)                                                           & \multicolumn{1}{c}{Ours} & Ours (EC2) & \multicolumn{1}{c}{Obliv-C}        \\ \midrule
-d=100                                                                  & 144                      & 243        & \textgreater $10^4$ \\ \midrule
-d=500                                                                  & 527                      & 2701       & \textgreater $10^3$ \\ \midrule
-\multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}SGD: \\ MNIST\end{tabular}} & \multicolumn{1}{c}{Ours} & \multicolumn{2}{c}{SecureML}                    \\ \cmidrule(l){2-4} 
-                                                                       & 10200                    & \multicolumn{2}{c}{10330}                       \\ \bottomrule
-\end{tabular}
-\end{table}
+When comparing runtimes, we found that SPDZ outperformed Obliv-C even in a distributed machine setting by an order of magnitude and comparable results to SecureML in SGD and Logistic Regression.
+The accuracy findings were comparable for LDLT, Cholesky, and CGD results for SPDZ and Obliv-C, but we found that SPDZ outperformed Obliv-C on SGD.
+SPDZ was able to match SecureML results on SGD and Logistic Regression.
 
-Through this work, we demonstrated that SPDZ outperforms other protocols Obliv-C and SecureML in terms of accuracy, as evidenced in Table 2, efficiency, and security.
-In future work, we plan to expand the selected algorithms beyond linear and logistic regression to include neural networks and deep neural networks.
+\section{Conclusion}
+
+Through this work, we demonstrated that SPDZ, a malicious security framework, can match the efficiency and accuracy other protocols of semi-honest through experiments on real and synthetic data.
 This investigation into the applicability of secure MPC on real world data demonstrates its potential change the way data is used in a variety of applications in today\textquotesingle s world.
@@
 
-[6] Mohassel, P., \& Zhang, Y. (2017). SecureML: A system for scalable privacy-preserving machine learning. {\it 38th IEEE Symposium on Security and Privacy (SP)}, pp. 19-38. IEEE.
+[6] Lindell, Y., \& Pinkas, B. (2009). A proof of security of Yao?s protocol for two-party computation. Journal of Cryptology, 22(2), 161-188.
 
-[7] Zahur, S., \& Evans, D. (2015). Obliv-C: A Language for Extensible Data-Oblivious Computation. {\it IACR Cryptology ePrint Archive, 2015}, 1153.
+[7] Mohassel, P., \& Zhang, Y. (2017). SecureML: A system for scalable privacy-preserving machine learning. {\it 38th IEEE Symposium on Security and Privacy (SP)}, pp. 19-38. IEEE.
+
+[8] Phung, D., Tseng, V.
+S., Webb, G.
+I., Ho, B., Ganji, M., \& Rashidi, L.
+(Eds.).
+(2018).
+Advances in Knowledge Discovery and Data Mining: 22nd Pacific-Asia Conference, PAKDD 2018, Melbourne, VIC, Australia, June 3-6, 2018, Proceedings (Vol.
+10938).
+Springer.
+
+[9] Zahur, S., \& Evans, D. (2015). Obliv-C: A Language for Extensible Data-Oblivious Computation. {\it IACR Cryptology ePrint Archive, 2015}, 1153.
 
 \end{document}
+
+%Can add table back in - I am describing results instead. 
+%\begin{table}[h!]
+%\caption{Runtime results against }
+%\centering
+%\begin{tabular}{@{}llcl@{}}
+%\toprule
+%\multicolumn{1}{c}{Datasets}                                           & \multicolumn{3}{c}{Runtime (seconds)}                                      \\ \midrule
+%CGD (64 bit)                                                           & \multicolumn{1}{c}{Ours} & Ours (EC2) & \multicolumn{1}{c}{Obliv-C}        \\ \midrule
+%d=100                                                                  & 144                      & 243        & \textgreater $10^3$ \\ \midrule
+%d=500                                                                  & 527                      & 2701       & \textgreater $10^4$ \\ \midrule
+%\multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}SGD: \\ MNIST\end{tabular}} & \multicolumn{1}{c}{Ours} & \multicolumn{2}{c}{SecureML}                    \\ \cmidrule(l){2-4} 
+%                                                                       & 10200                    & \multicolumn{2}{c}{10330}                       \\ \bottomrule
+%\end{tabular}
+%\end{table}

### 2018-10-13 paper/nips_format.tex
added revisions 2

@@
   Yale University\\
-  \textt{valerio.pastro@yale.edu}\\
+  \texttt{valerio.pastro@yale.edu}\\
   %% examples of more authors
@@
 Secure Multi-Party Computation (MPC) is an area of cryptography research that has the power to revolutionize computation between shared parties holding private data.
-However theoretical protocols are often not scalable and efficient given large quantities of real world data.
-This project investigates an MPC protocol called SPDZ which is secure against malicious adversaries in comparison to two other semi-honest MPC protocols Obliv-C and SecureML in machine learning and regression applications.
+However, theoretical protocols are often not scalable and inefficient when given large quantities of real world data.
+This project investigates an MPC protocol called SPDZ, which is secure against malicious adversaries, in comparison to two other semi-honest MPC protocols Obliv-C and SecureML in machine learning and regression applications.
 We demonstrate its potential to be used in real world applications with experiments to show that it outperforms the other two methods while maintaining a more rigorous level of security.
@@
 
-Regression analysis and other machine learning techniques aim to build a model that fits a set of predictors to the dependent variable.
-Models built from a large set of collected data can be used to predict future behavior.
+Regression analysis and other machine learning techniques aim to build a model that fits a set of predictors to a dependent variable.
 With the advent of big data, it is becoming increasingly common for entities to store vast amounts of user data, run regression analyses and build models of this data to gain insights.
-A desired use case could be, for example, the Center for Disease Control is interested in identifying disease outbreak and many individual hospitals have their own patient data which could benefit this study.
+A desired use case is computation on joint data from different private sources.
+For example, the Center for Disease Control is interested in identifying disease outbreak and many individual hospitals have their own patient data which could benefit this study.
 The problem is that openly sharing this data for prediction or model-building purposes would be against modern day privacy laws as it would leak private individual data [2].
@@
 
-Secure MPC addresses this problem by providing a mechanism through which different parties supply their data for joint computation without revealing individual values from each database or even intermediate values.
+Secure MPC addresses this problem by providing a mechanism through which different parties run jointly run a protocol on each of their private data for joint computation without revealing individual values or even intermediate values from each database and without the need for a trusted third party.
 This need for MPC motivates our study to bridge theory and application through experiments on real-world data.
 
-In general, MPC protocols predominantly come in two forms: security against semi-honest [6] and malicious adversaries [4].
-In semi-honest security, a more naive model, it is assumed that the party will follow the protocol as specified but will merely gather information intermediate values.
-Typically semi-honest security is more efficient on computation as it has fewer requirements.
+In general, MPC protocols come in two forms: security against semi-honest [6] and malicious adversaries [5].
+In semi-honest security, it is assumed that the party will follow the protocol as specified, but will merely gather information about intermediate values.
+Semi-honest security is more efficient on computation as it has fewer requirements than malicious security.
 In malicious security, dishonest parties may attempt to deviate from the specified protocol, and the protocol must be able to detect such cheating and abort the computation - thus any successful output is guaranteed to have privacy preserved.
@@
 
-For this study, we chose the SPDZ framework [1,5], which ensures against malicious security, and compare against two semi-honest methods Obliv-C [9] and SecureML [7].
-SPDZ performs the checks necessary to detect dishonest parties through using random data in the hopes of making computation more efficient.
-Obliv-C is based on Yao\textquotesingle s garbled circuits and [3] introduces optimizations on inner products and algorithmic modifications to speed up regression calculation.
-SecureML\textquotesingle s focus is with developing MPC-friendly alternatives to non-linear activation functions which are often intensive to compute as well as vectorization techniques in a shared setting.
+For this study, we chose the SPDZ framework [1,6], which ensures against malicious security, and compare against two semi-honest methods Obliv-C [10] and SecureML [8].
+The SPDZ protocol is comprised of an efficient preprocessing offline phase using semi-homomorphic encryption and an optimized secret-sharing online phase that computes an arithmetic circuit.
+Obliv-C is based on Yao\textquotesingle s garbled circuits and Gascon et al.[4]  introduces optimizations on inner products and algorithmic modifications to speed up regression calculation.
+SecureML developed MPC-friendly alternatives to non-linear activation functions which are often intensive to compute as well as vectorization techniques in a shared setting.
 
-Since secure computation slower and could be limited in accuracy when compared to computation in plaintext, there are a number of modifications to plaintext algorithms that can be made to obtain a reasonable degree of accuracy in more efficient time.
+Secure computation is slower and could have lower accuracy when compared to the same computation in plaintext.
+Thus trade-offs need to be made in terms of the accuracy and the complexity of the algorithm, which will affect the MPC efficiency.
 
@@
 
-Our goal was to conduct a thorough investigation of the SPDZ framework in the applications of regression and machine learning algorithms in a similar fashion to that done in  [3,6], evaluating in terms of runtime and accuracy.
-The main algorithms that we considered were, LDLT Decomposition, Cholesky, Conjugate Gradient Descent (CGD), Stochastic Gradient Descent (SGD), Logistic Regression.
-The algorithms were implemented in the SPDZ framework as a proof-of-concept as well as in python as a plaintext verification of the accuracy of the secure version.
+Our goal was to evaluate and compare the performance of SPDZ against other frameworks in the context of ML applications that have been previously considered in [4,8], evaluating in terms of runtime and accuracy.
+The main algorithms that we considered were: LDLT Decomposition, Cholesky, Conjugate Gradient Descent (CGD), Stochastic Gradient Descent (SGD), Logistic Regression.
+These are classic algorithms that can be found in standard machine textbooks (e.g., [3]).
+In our evaluation, the algorithms were implemented in the SPDZ framework as well as in python as a plaintext verification of the algorithm.
 
@@
 The algorithms are cubic in complexity with asymptotic run time of $O(d^3)$, where $d$ is the dimension of the input matrix.
-This can become inefficient to work with when looking at real world data that has large number of features and entries.
+This runtime can become inefficient when considering real world data that have large number of features and entries.
 The difference between LDLT and Cholesky is that Cholesky requires a square root.
-In MPC, non-linear functions like square roots are difficult to represent.
-Thus we looked at an iterative Newton method as a way of approximating the square root in SPDZ to solve for $x_i$ where $x_{i}^2 = S$:
+In MPC, non-linear functions like square roots are difficult to represent due to the linear nature of secret sharing schemes.
+Thus, we looked at an iterative Newton method as a way of approximating the square root in SPDZ to solve for $x_i$, where $x_{i}^2 = S$:
 
@@
 
-In terms of an iterative approach to regression, [3] proposed a modification for the normalization in the iterative version of CGD that would be more stable with fixed-point numbers.
+In terms of an iterative approach to regression, Gascon et al.[4] proposed a modification for the normalization in the iterative version of CGD that would be more stable with fixed-point numbers.
 We applied the same algorithm to SPDZ to see if the same modification would work.
@@
 
-Stochastic gradient descent is also an iterative method that minimizes a differentiable cost function and is the driving mechanism behind most neural networks.
-In linear regression the cost function is derived from the mean squared error.
-The update consists of forward propagation to calculate the predicted output $y_i^* = x_i \cdot w$ and a backward propagation phase to adjust weights accordingly given differential error and learning rate $\alpha(y_i^*-y_i)x_{ij}$ for an update function of $w_j = w_j  - \alpha(x_i \cdot w *-y_i)x_{ij}$, where $w$ are the weights, $\alpha$ the learning rate, $x$ the input, and $y$ the expected output [7].
+Stochastic gradient descent is an iterative approximation method that converges to the global minimum for convex problems, like linear regression.
+It is also a driving mechanism for non-convex problems like neural networks.
+In a linear regression setting, we update the learned weights $w$, given learning rate $\alpha$ and input matrix $x$ and expected output $y$, using $w_j = w_j  - \alpha(x_i \cdot w *-y_i)x_{ij}$ [8].
+In this update function, the weights are adjusted element-wise by the error from the predicted and expected value at a rate determined by $\alpha$.
 
-We implemented mini-batch SGD so the weights would converge more quickly and in smoother fashion to the minimum.
 Instead of selecting one sample per iteration, a small batch of size $B$ samples are selected and the update function is performed collectively.
-Since SPDZ had limited functionality for matrix manipulations and vectorization, we were not able to take advantage of the time speed-ups in this sense, but were still able to take advantage of the accuracy benefit.
-The update function for the batched version, where $w$ are the weights, $\alpha$ the learning rate, $X_{B}$ the batched input, and $Y_{B}$ the corresponding expected batched output:
+We implemented this modification called mini-batch SGD so the weights would converge more quickly and in smoother fashion to the minimum.
+Since SPDZ does not support vectorization (e.g.
+in python), where operations are broadcasted to entire array, we were not able to take advantage of the time speed-ups.
+However we were still able to take advantage of the accuracy benefit.
+The algorithm for the batched version is modified as such, where $w$ are the weights, $\alpha$ the learning rate, $X_{B}$ the batched input, and $Y_{B}$ the corresponding expected batched output:
 \[
@@
 
-Logistic regression models a binary dependent variable.
-For logistic regression, the activation function is defined as the logistic function $f(u) = \frac{1}{1+e^-u}$.
-Using the cross entropy function, we derive a similar update function for mini-batched SGD as $ w = w - \frac{1}{|B|} \alpha X^{T}_{B} \times f(X_{B} \times w - Y_{B})$.
-To approximate the activation function, [2] proposed this following piecewise function:
+Logistic regression is a classification algorithm for modeling a binary dependent variable.
+Logistic regression fits the logistic function $f(u) = \frac{1}{1+e^-u}$ to the input.
+Using the cross entropy function, we can derive an update function for mini-batched SGD in a similar form to linear regression: $ w = w - \frac{1}{|B|} \alpha X^{T}_{B} \times f(X_{B} \times w - Y_{B})$, where $f$ maps the predicted value into the binary output space.
+To approximate this function.
+$f$, Mohassel and Zhang [8] proposed this following piecewise function:
 
@@
 
-The main metrics of evaluations were the latency in evaluation in a privacy-preserving setting and the error in accuracy, to see what trade-offs and modifications should be made to achieve good accuracy while optimizing efficiency.
+The main metrics of evaluations were the latency in computation in a privacy-preserving setting and the error in accuracy, to see what trade-offs and modifications should be made to achieve good accuracy while optimizing efficiency.
 We varied the precision after the decimal point (32 and 64 bits) or less depending on the comparison.
 
-To evaluate the SPDZ protocol implementation, we needed to identify a source of data.
-We used both a combination of existing, well-known datasets (MNIST, Arcene, and 9 other UCI open-source datasets) as well as synthetically generated data.
-The advantage to using synthetic data is being able to control characteristics of the data like the dimensions ($d$ = 10, 20, 50, 100, 200, 500), condition number ($cd$ = [1,10]), and number of examples ($n$ = 1000, 100000).
+We evaluated our methods on both real-world datasets (MNIST, Arcene, and 9 other UCI open-source datasets) as well as synthetically generated data.
+These real-world datasets allow us to compare the accuracy results to existing works and to demonstrate that SPDZ can be used in practical settings.
+However, the advantage to using synthetic data is being able to control characteristics of the data like the dimensions ($d$ = 10, 20, 50, 100, 200, 500), condition number ($cd$ = [1,10]), and number of examples ($n$ = 1000, 100000).
 
@@
 The larger the condition number, the larger the error in approximations of the solution.
-The direct decomposition methods grew exponentially in run time on the log scale as the matrix size increased, which means that it is not suitable for large size real data.
+The direct decomposition methods grew exponentially in run time as input size increases, which is shown in the left column of Figure 1, which means that it is not suitable for large size real data.
 Alternatively, the iterative CGD runtime increases at a much slower rate.
@@
   \vspace{-4mm}
-   \caption{Comparing accuracy of privacy preserving linear regression with fixed point and plaintext training on floating point for MNIST (left) and Arcene (right) in [2].}
+   \caption{Comparing accuracy of privacy preserving linear regression with various fixed point precisions and plaintext training on floating point for MNIST (left) and Arcene (right).}
    \vspace{-4mm}
@@
 
-Figure 2, which compares MNIST and Arcene results, demonstrates that the number of bits of precision needed to get good accuracy is highly dependent on the dataset.
+Figure 2, which compares SGD on MNIST and Arcene results, demonstrates that the number of bits of precision needed to get good accuracy is highly dependent on the dataset.
 For MNIST, 13 bits was sufficient to match plaintext accuracy but 28 bits were needed for Arcene.
 The MNIST data contains only 784 features while there are 10,000 in the Arcene data, 3,000 of which are considered "probes" with no predictive power, which could explain the lower overall accuracy [8].
-While the numbers in the MNIST data ranged from 0 to 9, to match the approach in [7], we only used the 0s and 1s labels from the dataset, thus reducing it to a binary problem.
-We did run the computation with all 10 digits, but found that gradient descent achieved a much lower accuracy of about 19\%.
-We also compared the root mean squared error (RMSE) of SGD on 9 UCI open-sourced datasets of ranging sizes to results in [3] and found lower RMSE results.
-Due to limited space, these results are not presented.
+While the numbers in the MNIST data ranged from 0 to 9, Mohassel and Zhang [8] only used 0s and 1s labels from the dataset, reducing it to a binary problem.
+We replicated this approach and present the results below.
+We did run the computation to predict all 10 digits, but found that SGD only achieved a much lower accuracy of about 19\%.
+We also compared the root mean squared error (RMSE) of SGD on 9 UCI open-sourced datasets of ranging sizes to results in [3].
+Our results in the SPDZ secure settings typically increased RMSE by about $5-20\%$ compared to plaintext computation, but still outperformed RMSE results from [3] in both CGD and SGD.
 
@@
 
-\subsection{Comparison}
+\subsection{Conclusion}
 
@@
 SPDZ was able to match SecureML results on SGD and Logistic Regression.
-
-\section{Conclusion}
-
-Through this work, we demonstrated that SPDZ, a malicious security framework, can match the efficiency and accuracy other protocols of semi-honest through experiments on real and synthetic data.
-This investigation into the applicability of secure MPC on real world data demonstrates its potential change the way data is used in a variety of applications in today\textquotesingle s world.
 
@@
 
-[3] Gasc\'{o}n, A., Schoppmann, P., Balle, B., Raykova, M., Doerner, J., Zahur, S., \& Evans, D.
+[3] Friedman, J., Hastie, T., \& Tibshirani, R. (2001). The elements of statistical learning (Vol. 1, No. 10). New York, NY, USA:: Springer series in statistics.
+
+[4] Gasc\'{o}n, A., Schoppmann, P., Balle, B., Raykova, M., Doerner, J., Zahur, S., \& Evans, D.
 (2017).
@@
 
-[4] Hazay, C., \& Lindell, Y. (2010). A Note on the Relation between the Definitions of Security for Semi-Honest and Malicious Adversaries. {\it IACR Cryptology ePrint Archive}, pp.\ 551.
+[5] Hazay, C., \& Lindell, Y. (2010). A Note on the Relation between the Definitions of Security for Semi-Honest and Malicious Adversaries. {\it IACR Cryptology ePrint Archive}, pp.\ 551.
 
-[5] Keller, M., Scholl, P., \& Smart, N.
+[6] Keller, M., Scholl, P., \& Smart, N.
 P.
@@
 
-[6] Lindell, Y., \& Pinkas, B. (2009). A proof of security of Yao?s protocol for two-party computation. Journal of Cryptology, 22(2), 161-188.
+[7] Lindell, Y., \& Pinkas, B. (2009). A proof of security of Yao?s protocol for two-party computation. Journal of Cryptology, 22(2), 161-188.
 
-[7] Mohassel, P., \& Zhang, Y. (2017). SecureML: A system for scalable privacy-preserving machine learning. {\it 38th IEEE Symposium on Security and Privacy (SP)}, pp. 19-38. IEEE.
+[8] Mohassel, P., \& Zhang, Y. (2017). SecureML: A system for scalable privacy-preserving machine learning. {\it 38th IEEE Symposium on Security and Privacy (SP)}, pp. 19-38. IEEE.
 
-[8] Phung, D., Tseng, V.
+[9] Phung, D., Tseng, V.
 S., Webb, G.
@@
 
-[9] Zahur, S., \& Evans, D. (2015). Obliv-C: A Language for Extensible Data-Oblivious Computation. {\it IACR Cryptology ePrint Archive, 2015}, 1153.
+[10] Zahur, S., \& Evans, D. (2015). Obliv-C: A Language for Extensible Data-Oblivious Computation. {\it IACR Cryptology ePrint Archive, 2015}, 1153.
 
@@
 %\end{table}
+
+%\section{Conclusion}
+
+%Through this work, we demonstrated that SPDZ, a malicious security framework, can match the efficiency and accuracy other protocols of semi-honest through experiments on real and synthetic data.
+This investigation into the applicability of secure MPC on real world data demonstrates its potential change the way data is used in a variety of applications in today\textquotesingle s world.

### 2018-10-14 paper/nips_format.tex
edits

@@
 \begin{abstract}
-Secure Multi-Party Computation (MPC) is an area of cryptography research that has the power to revolutionize computation between shared parties holding private data.
-However, theoretical protocols are often not scalable and inefficient when given large quantities of real world data.
-This project investigates an MPC protocol called SPDZ, which is secure against malicious adversaries, in comparison to two other semi-honest MPC protocols Obliv-C and SecureML in machine learning and regression applications.
-We demonstrate its potential to be used in real world applications with experiments to show that it outperforms the other two methods while maintaining a more rigorous level of security.
+Secure Multi-Party Computation (MPC) is an area of cryptography, which enables the computation on sensitive data from multiple sources while maintaining privacy guarantees.
+However, theoretical protocols often do not scalable efficiently to real world size data. This project investigates the efficiency of of the SPDZ framework~\cite{SPDZ12}, which provides implementation
+of MPC protocols with malicious security, in the context of popular machine learning (ML) algorithms. In particular, we choose applications such as linear regression and logistic regression,
+which have been implemented and evaluated using semi-honest MPC techniques~\cite{GSB0DZE17, MZ17}.
+We demonstrate that the SPDZ framework outperforms these previous implementations while providing stronger security.
 \end{abstract}
@@
 
-Regression analysis and other machine learning techniques aim to build a model that fits a set of predictors to a dependent variable.
-With the advent of big data, it is becoming increasingly common for entities to store vast amounts of user data, run regression analyses and build models of this data to gain insights.
-A desired use case is computation on joint data from different private sources.
-For example, the Center for Disease Control is interested in identifying disease outbreak and many individual hospitals have their own patient data which could benefit this study.
-The problem is that openly sharing this data for prediction or model-building purposes would be against modern day privacy laws as it would leak private individual data [2].
-This application of secure MPC is one that would be extremely valuable in today\textquotesingle s world.
+Many machine learning techniques, including regression analysis, aim to build a model that fits a set of predictors to a dependent variable.
+Such techniques are widely used to model and analyze big data.
+In many settings, however, the input data for such ML analysis tools is partitioned among different parties, which have strict privacy policies.
+For example, the Center for Disease Control is interested in identifying disease outbreak and many individual hospitals have their own patients' data which could benefit this study.
+The problem is that openly sharing this data for prediction or model-building purposes would be against modern day privacy laws as it would leak private individual data.
+This is one of many examples, which could benefit of the functionality that allows to evaluate to output of the analysis without revealing more about the private inputs.
+%This application of secure MPC is one that would be extremely valuable in today\textquotesingle s world.
 
@@
 
-Secure MPC addresses this problem by providing a mechanism through which different parties run jointly run a protocol on each of their private data for joint computation without revealing individual values or even intermediate values from each database and without the need for a trusted third party.
-This need for MPC motivates our study to bridge theory and application through experiments on real-world data.
+Secure MPC addresses the above problem by providing a mechanism through which different parties can run a joint computation over their private inputs with guarantees that the only thing
+revealed about the inputs, is the output of the computation and whatever can be inherently inferred from it.
+There are two main types of MPC protocols in terms of their security guarantees: semi-honest and malicious protocols \cite{Goldreich:2004:FCV:975541, cryptoeprint:2010:551}.
+In semi-honest security, it is assumed that the parties will follow the protocol as specified, but they can try to infer information about the input from the protocol messages.
+In malicious security, dishonest parties may attempt to deviate from the specified protocol, and the protocol must guarantee that these parties cannot learn about the inputs.
+Since malicious protocols have to satisfy stronger guarantees in general such construction are less efficient than semi-honest protocols.
 
-In general, MPC protocols come in two forms: security against semi-honest [6] and malicious adversaries [5].
-In semi-honest security, it is assumed that the party will follow the protocol as specified, but will merely gather information about intermediate values.
-Semi-honest security is more efficient on computation as it has fewer requirements than malicious security.
-In malicious security, dishonest parties may attempt to deviate from the specified protocol, and the protocol must be able to detect such cheating and abort the computation - thus any successful output is guaranteed to have privacy preserved.
-This notion can be rigorously proved.
+Two recent works~\cite{GSB0DZE17, MZ17} propose efficient implementation for several central building blocks for machine learning such as conjugate gradient decent (CGD) and stochastic gradient descent (SGD) as well as applications  using them such as linear and logistic regression.
+The work of Gascon et al.~\cite{GSB0DZE17} uses the framework for semi-honest computation Obliv-C~\cite{cryptoeprint:2015:1153} and proposes
+several different methods for solving systems of linear equations.
+Their main premise is to use an iterative method such as CGD and show trade-offs that save a lot of computation and hence efficiency for the MPC in
+return for a small accuracy loss.
+They further propose a modification of CGD that has stable behavior using fix-point arithmetic since emulating floating point with the underlying MPC representation introduces substantial
+efficiency overhead.
 
+The work of Mohassel and Zhang~\cite{MZ17} considers stochastic gradient descent as a method for learning linear regression and logistic regression by using different activation functions.
+The authors consider
+arithmetic representation for the computation and propose new secure computation techniques for matrix computation, which generalize the approach for generating multiplicative triples in a preprocessing step.
+ Similarly to the work of Gascon et al.~\cite{GSB0DZE17}, this paper considers techniques for approximation that save computation, for example, using a piece-wise approximation for the logistic function.
+The author
+ also propose new techniques for more efficient approximate computation of fixed-point encodings.
+
+The techniques in both of the above works are restricted to the setting of two party computation.
+We choose the SPDZ~\cite{SPDZ12,KSS13,KPR18} framework for our implementations since it is one of the main and most comprehensive
+implementations for multiparty computation protocols, which provide malicious security and support more than two parties.
+
+\iffalse
 For this study, we chose the SPDZ framework [1,6], which ensures against malicious security, and compare against two semi-honest methods Obliv-C [10] and SecureML [8].
@@
 Thus trade-offs need to be made in terms of the accuracy and the complexity of the algorithm, which will affect the MPC efficiency.
+\fi
 
@@
 
-Our goal was to evaluate and compare the performance of SPDZ against other frameworks in the context of ML applications that have been previously considered in [4,8], evaluating in terms of runtime and accuracy.
-The main algorithms that we considered were: LDLT Decomposition, Cholesky, Conjugate Gradient Descent (CGD), Stochastic Gradient Descent (SGD), Logistic Regression.
+%Our goal was to evaluate and compare the performance of SPDZ against other frameworks in the context of ML applications that have been previously considered in [4,8], evaluating in terms of runtime and accuracy.
+For our implementation we consider the same algorithms and functionalities as in the above two papers.
+Next we provide a brief overview of these classic algorithms (details can be found in ~\cite{HTF09}).
+
+%The main algorithms that we considered were: LDLT Decomposition, Cholesky, Conjugate Gradient Descent (CGD), Stochastic Gradient Descent (SGD), Logistic Regression.
 These are classic algorithms that can be found in standard machine textbooks (e.g., [3]).
-In our evaluation, the algorithms were implemented in the SPDZ framework as well as in python as a plaintext verification of the algorithm.
 
-\subsection{Direct vs. Iterative Decomposition}
+%In our evaluation, the algorithms were implemented in the SPDZ framework as well as in python as a plaintext verification of the algorithm.
 
+\subsection{Direct vs. Iterative Decomposition for Solving a Linear System}
+
+Solving a system of linear equations, which underlies linear regression learning, can be done using techniques for direct and indirect decomposition.
 LDLT and Cholesky are both variants of direct decomposition methods which decompose a Hermitian, positive-definite matrix into a lower triangular matrix and its conjugate transpose.
 The algorithms are cubic in complexity with asymptotic run time of $O(d^3)$, where $d$ is the dimension of the input matrix.
-This runtime can become inefficient when considering real world data that have large number of features and entries.
-The difference between LDLT and Cholesky is that Cholesky requires a square root.
+%This runtime can become inefficient when considering real world data that have large number of features and entries. 
+The difference between LDLT and Cholesky is that Cholesky requires a square root. The representation of square root computation as an arithmetic circuit used in the MPC computation in SPDZ introduces 
+considerable overhear. That is why we used the iterative Newton method as a way of approximating the square root computation. It computes $x_i$, where $x_{i}^2 = S$, with repetition of the following
+update function $x_{n+1} =1/2 (x_{n} + S/x_{n})$.
+
+\iffalse
 In MPC, non-linear functions like square roots are difficult to represent due to the linear nature of secret sharing schemes.
@@
 \]
+\fi
 
-In terms of an iterative approach to regression, Gascon et al.[4] proposed a modification for the normalization in the iterative version of CGD that would be more stable with fixed-point numbers.
+In terms of an iterative approach to regression, we used the approach proposed by
+Gascon et al.~\cite{GSB0DZE17}, which uses a normalized version of CGD that preserves stability and convergence rate with fixed-point number representation. Similarly to other MPC implementations
+using floating point representation in SPDZ introduces substantial efficiency overhead.
+
+%proposed a modification for the normalization in the iterative version of CGD that would be more stable with fixed-point numbers.
 We applied the same algorithm to SPDZ to see if the same modification would work.
@@
 
-\subsection{Gradient Descent}
+\subsection{Stochastic Gradient Descent}
 
-Stochastic gradient descent is an iterative approximation method that converges to the global minimum for convex problems, like linear regression.
+Stochastic gradient descent is an iterative approximation method that converges to the global minimum for convex problems, like linear and logistic regression.
 It is also a driving mechanism for non-convex problems like neural networks.
+An SGD iteration updates a weight vector $\mathbf{w}$ using a randomly selected sample from the training input as follows: $w_j := w_j - \alpha ({\partial C_i(\mathbf{w})}/{\partial w_j})$ with learning rate $\alpha$.
+In this update $C_i$ is the cost function, which can be instantiated with different concrete functions to obtain computation for linear regression and logistic regression.
+A common technique for SGD
+computation is called \emph{mini-batch} -- instead of selecting one sample per iteration, a small batch of size $B$ samples are selected and the update function is performed averaging the partial derivatives across all samples.
+We use the mini-batch SGD in our implementation to obtain accuracy benefits.
+While the work of Mohassel and Zhang~\cite{MZ17} has optimizations for matrix computation, which can be used with mini-batch, for
+SPDZ this does not lead to additional savings.
+
+\iffalse
 In a linear regression setting, we update the learned weights $w$, given learning rate $\alpha$ and input matrix $x$ and expected output $y$, using $w_j = w_j  - \alpha(x_i \cdot w *-y_i)x_{ij}$ [8].
@@
 \]
+\fi
 
-\subsubsection{Non-Linear Activations}
+\subsubsection{Linear and Non-Linear Activation Functions}
+
+To obtain a solution for linear regression using SGD, we instantiate the cost function as $w_j = w_j  - \alpha(X_i \cdot w *-y_i)X_{ij}$, where $X$ is the input matrix and $y$ is the input vector.
+In this update function, the weights are adjusted element-wise by the error from the predicted and expected value at a rate determined by $\alpha$.
 
@@
 Logistic regression fits the logistic function $f(u) = \frac{1}{1+e^-u}$ to the input.
-Using the cross entropy function, we can derive an update function for mini-batched SGD in a similar form to linear regression: $ w = w - \frac{1}{|B|} \alpha X^{T}_{B} \times f(X_{B} \times w - Y_{B})$, where $f$ maps the predicted value into the binary output space.
-To approximate this function.
-$f$, Mohassel and Zhang [8] proposed this following piecewise function:
+The corresponding update function for mini-batched SGD for logistic regression is $ w = w - \frac{1}{|B|} \alpha X^{T}_{B} \times f(X_{B} \times w - Y_{B})$, where $f$ maps the predicted value into the binary output space.
+Mohassel and Zhang [8] proposed the following piecewise function as approximation for $f$:
 
@@
 
-The main metrics of evaluations were the latency in computation in a privacy-preserving setting and the error in accuracy, to see what trade-offs and modifications should be made to achieve good accuracy while optimizing efficiency.
-We varied the precision after the decimal point (32 and 64 bits) or less depending on the comparison.
+For our evaluation, we implemented all algorithms both in the SPDZ framework as well as in python as a plaintext verification of the algorithm.
+The main metrics of evaluations were the latency of the MPC computation and the accuracy error, and we aimed to explore the trade-offs between accuracy and efficiency.
+We varied the precision after the decimal point
+depending on what was used in the works that we compared against (32 and 64 bits for the linear regression, less for SGD).
 
@@
 These real-world datasets allow us to compare the accuracy results to existing works and to demonstrate that SPDZ can be used in practical settings.
-However, the advantage to using synthetic data is being able to control characteristics of the data like the dimensions ($d$ = 10, 20, 50, 100, 200, 500), condition number ($cd$ = [1,10]), and number of examples ($n$ = 1000, 100000).
+We used synthetic data in order to explore larger ranges of data characteristics such as dimension ($d$ = 10, 20, 50, 100, 200, 500), condition number ($cd$ = [1,10]), and number of examples ($n$ = 1000, 100000).
 
-In our experiments, both parties were mostly run on one machine to simplify the procedure, but we also performed tests where both parties were deployed on separate Amazon EC2 m4.large instances.
+Most of our experiments were ran using machines on the same local area network where there is no network latency.
+We performed tests where both parties were deployed on separate Amazon EC2 m4.large instances.
+We also ran experiment with up to four parties.
+
+%In our experiments, both parties were mostly run on one machine to simplify the procedure, but we also performed tests where both parties were deployed on separate Amazon EC2 m4.large instances.
 It is important to note that the distribution of players across machine only affects runtime and not the accuracy of the results.
@@
 
-This section presents empirical results for real-world and synthetic data evaluated on the SPDZ protocol, comparing the five different algorithms in terms of accuracy and run time for various parameters.
+In this section we present empirical results for our SPDZ implementations evaluated with real and synthetic databases.
+We compare the five different algorithms in terms of accuracy and run time for various parameters.
 
@@
   \includegraphics[scale=0.4]{allregression.png}
-  \label{fig:result4}
     \caption{(Left) Run time as a function of input dimension.
@@
 (Bottom) Fixed-point with 28 bits of precision}
+   \label{fig:result1}   
 \end{figure}
@@
 For LDLT, Cholesky and various iterations of CGD, we evaluated on synthetically generated data of varying sizes and condition numbers.
-The larger the condition number, the larger the error in approximations of the solution.
-The direct decomposition methods grew exponentially in run time as input size increases, which is shown in the left column of Figure 1, which means that it is not suitable for large size real data.
+The larger the condition number is, the larger the error in approximations of the solution is.
+The direct decomposition methods grew exponentially in run time as input size increases, which is shown in the left column of Figure~\ref{fig:result1} -- this unlikely to be suitable for large size real data.
 Alternatively, the iterative CGD runtime increases at a much slower rate.
-In the middle column of Figure 2, we find that about 20 iterations is sufficient to reach maximum accuracy given the number of allocated bits even with varying condition numbers, meaning the data might not be as well-formed.
+In the middle column of Figure 2, we find that about 20 iterations are sufficient to reach maximum accuracy given the number of allocated bits even with varying condition numbers, meaning the data might not be as well-formed.
 Particularly for the 64-bit case, shown on the bottom right, the accuracy is identical for CGD after 15 iterations and Cholesky/LDLT.
@@
   \includegraphics[scale=0.6]{mnistarcene.png}
-  \label{fig:result4}
   \vspace{-4mm}
    \caption{Comparing accuracy of privacy preserving linear regression with various fixed point precisions and plaintext training on floating point for MNIST (left) and Arcene (right).}
-   \vspace{-4mm}
+   %\vspace{-4mm}
+   \label{fig:result2}  
 \end{figure}
 
-Figure 2, which compares SGD on MNIST and Arcene results, demonstrates that the number of bits of precision needed to get good accuracy is highly dependent on the dataset.
+Figure~\ref{fig:result2} compares SGD on MNIST and Arcene results.
+It shows that the number of bits of precision needed to get good accuracy is highly dependent on the dataset.
 For MNIST, 13 bits was sufficient to match plaintext accuracy but 28 bits were needed for Arcene.
 The MNIST data contains only 784 features while there are 10,000 in the Arcene data, 3,000 of which are considered "probes" with no predictive power, which could explain the lower overall accuracy [8].
-While the numbers in the MNIST data ranged from 0 to 9, Mohassel and Zhang [8] only used 0s and 1s labels from the dataset, reducing it to a binary problem.
+While the numbers in the MNIST data ranged from 0 to 9, Mohassel and Zhang~\cite{MZ17} only used 0s and 1s labels from the dataset, reducing it to a binary problem.
 We replicated this approach and present the results below.
@@
 We also compared the root mean squared error (RMSE) of SGD on 9 UCI open-sourced datasets of ranging sizes to results in [3].
-Our results in the SPDZ secure settings typically increased RMSE by about $5-20\%$ compared to plaintext computation, but still outperformed RMSE results from [3] in both CGD and SGD.
+Our results in the SPDZ secure setting typically increased RMSE by about $5-20\%$ compared to plaintext computation, but still outperformed RMSE results from [3] in both CGD and SGD.
 
-In terms of logistic regression, for SPDZ, we did not find that the new activation function was a better alternative to taking a Taylor Series approximation for the exponential function as shown in Table 1.
+In terms of logistic regression, for SPDZ, we did not find that the new activation function was a better alternative to taking a Taylor Series approximation for the exponential function as shown in Table~\ref{tbl1}.
 We found that for SPDZ, which is based on arithmetic circuits, the extra time to take a few extra degrees in the approximation was negligible.
@@
 \vspace{-4mm}
+\label{tbl1}
 \end{table}
@@
 
-When comparing runtimes, we found that SPDZ outperformed Obliv-C even in a distributed machine setting by an order of magnitude and comparable results to SecureML in SGD and Logistic Regression.
+We found that the SPDZ outperformed Obliv-C even in a distributed machine setting by an order of magnitude and achieves comparable results to SecureML for SGD instantiations such as Linear and Logistic Regression.
 The accuracy findings were comparable for LDLT, Cholesky, and CGD results for SPDZ and Obliv-C, but we found that SPDZ outperformed Obliv-C on SGD.
@@
 
-\section*{References}
+%\section*{References}
+
+\bibliographystyle{plain}
+\bibliography{refs}
+
+\iffalse
 
@@
 
+\fi
+
 \end{document}

### 2018-10-15 paper/nips_format.tex
edits

@@
 
-\title{Experimental Secure Multi-Party Computation on Real Data using SPDZ}
+\title{Secure Computation for Machine Learning With SPDZ}
 
@@
 Alternatively, the iterative CGD runtime increases at a much slower rate.
-In the middle column of Figure 2, we find that about 20 iterations are sufficient to reach maximum accuracy given the number of allocated bits even with varying condition numbers, meaning the data might not be as well-formed.
+In the middle column of Figure 2, we find that about 20 iterations are sufficient to reach maximum accuracy given the number of allocated bits even with varying condition numbers.
 Particularly for the 64-bit case, shown on the bottom right, the accuracy is identical for CGD after 15 iterations and Cholesky/LDLT.
@@
 For MNIST, 13 bits was sufficient to match plaintext accuracy but 28 bits were needed for Arcene.
-The MNIST data contains only 784 features while there are 10,000 in the Arcene data, 3,000 of which are considered "probes" with no predictive power, which could explain the lower overall accuracy [8].
+The MNIST data contains only 784 features while there are 10,000 in the Arcene data, 3,000 of which are considered "probes" with no predictive power, which could explain the lower overall accuracy of \cite{MZ17}.
 While the numbers in the MNIST data ranged from 0 to 9, Mohassel and Zhang~\cite{MZ17} only used 0s and 1s labels from the dataset, reducing it to a binary problem.
@@
 We did run the computation to predict all 10 digits, but found that SGD only achieved a much lower accuracy of about 19\%.
-We also compared the root mean squared error (RMSE) of SGD on 9 UCI open-sourced datasets of ranging sizes to results in [3].
-Our results in the SPDZ secure setting typically increased RMSE by about $5-20\%$ compared to plaintext computation, but still outperformed RMSE results from [3] in both CGD and SGD.
+We also compared the root mean squared error (RMSE) of SGD on 9 UCI open-sourced datasets of ranging sizes to results in \cite{GSB0DZE17}.
+Our results in the SPDZ secure setting typically increased RMSE by about $5-20\%$ compared to plaintext computation, but still outperformed RMSE results from \cite{GSB0DZE17} in both CGD and SGD.
 

### 2018-10-16 paper/nips_format.tex
minor text edits, added supplementary material

@@
 \begin{abstract}
-Secure Multi-Party Computation (MPC) is an area of cryptography, which enables the computation on sensitive data from multiple sources while maintaining privacy guarantees.
-However, theoretical protocols often do not scalable efficiently to real world size data. This project investigates the efficiency of of the SPDZ framework~\cite{SPDZ12}, which provides implementation
-of MPC protocols with malicious security, in the context of popular machine learning (ML) algorithms. In particular, we choose applications such as linear regression and logistic regression,
+Secure Multi-Party Computation (MPC) is an area of cryptography that enables the computation on sensitive data from multiple sources while maintaining privacy guarantees.
+However, theoretical MPC protocols often do not scale efficiently to real-world data. This project investigates the efficiency of of the SPDZ framework~\cite{SPDZ12}, which provides an implementation
+of an MPC protocol with malicious security, in the context of popular machine learning (ML) algorithms. In particular, we chose applications such as linear regression and logistic regression,
 which have been implemented and evaluated using semi-honest MPC techniques~\cite{GSB0DZE17, MZ17}.
@@
 In many settings, however, the input data for such ML analysis tools is partitioned among different parties, which have strict privacy policies.
-For example, the Center for Disease Control is interested in identifying disease outbreak and many individual hospitals have their own patients' data which could benefit this study.
-The problem is that openly sharing this data for prediction or model-building purposes would be against modern day privacy laws as it would leak private individual data.
-This is one of many examples, which could benefit of the functionality that allows to evaluate to output of the analysis without revealing more about the private inputs.
+For example, if the Center for Disease Control is interested in identifying disease outbreak, they might want to incorporate patient data from many individual hospitals.
+The problem is that openly sharing this data for prediction or model-building purposes is against modern day privacy laws as it would leak private individual data.
+This is one of many examples, which could benefit from MPC which would allow the parties to evaluate the output of the analysis without revealing more about the private inputs.
 %This application of secure MPC is one that would be extremely valuable in today\textquotesingle s world.
@@
 Secure MPC addresses the above problem by providing a mechanism through which different parties can run a joint computation over their private inputs with guarantees that the only thing
-revealed about the inputs, is the output of the computation and whatever can be inherently inferred from it.
+revealed about the inputs is the output of the computation and whatever can be inherently inferred from it.
 There are two main types of MPC protocols in terms of their security guarantees: semi-honest and malicious protocols \cite{Goldreich:2004:FCV:975541, cryptoeprint:2010:551}.
@@
 In malicious security, dishonest parties may attempt to deviate from the specified protocol, and the protocol must guarantee that these parties cannot learn about the inputs.
-Since malicious protocols have to satisfy stronger guarantees in general such construction are less efficient than semi-honest protocols.
+Since malicious protocols have to satisfy stronger guarantees in general, such construction are less efficient than semi-honest protocols.
 
@@
 several different methods for solving systems of linear equations.
-Their main premise is to use an iterative method such as CGD and show trade-offs that save a lot of computation and hence efficiency for the MPC in
+Their main premise is to use an iterative method such as CGD and demonstrate trade-offs that saves computation and hence efficiency for the MPC in
 return for a small accuracy loss.
-They further propose a modification of CGD that has stable behavior using fix-point arithmetic since emulating floating point with the underlying MPC representation introduces substantial
+They further propose a modification of CGD that has stable behavior using fixed-point arithmetic as emulating floating point with the underlying MPC representation introduces substantial
 efficiency overhead.
 
-The work of Mohassel and Zhang~\cite{MZ17} considers stochastic gradient descent as a method for learning linear regression and logistic regression by using different activation functions.
+The work of Mohassel and Zhang~\cite{MZ17} considers stochastic gradient descent as a method for linear regression and logistic regression with the incorporation of different activation functions.
 The authors consider
 arithmetic representation for the computation and propose new secure computation techniques for matrix computation, which generalize the approach for generating multiplicative triples in a preprocessing step.
- Similarly to the work of Gascon et al.~\cite{GSB0DZE17}, this paper considers techniques for approximation that save computation, for example, using a piece-wise approximation for the logistic function.
-The author
+ Similarly to the work of Gascon et al.~\cite{GSB0DZE17}, this paper considers techniques for approximation that save computation.
+For example, they use a piece-wise approximation of the logistic function.
+The authors
  also propose new techniques for more efficient approximate computation of fixed-point encodings.
@@
 The techniques in both of the above works are restricted to the setting of two party computation.
-We choose the SPDZ~\cite{SPDZ12,KSS13,KPR18} framework for our implementations since it is one of the main and most comprehensive
+We selected the SPDZ~\cite{SPDZ12,KPR18,KSS13} framework for our experiments since it is one of the main and most comprehensive
 implementations for multiparty computation protocols, which provide malicious security and support more than two parties.
@@
 
-To obtain a solution for linear regression using SGD, we instantiate the cost function as $w_j = w_j  - \alpha(X_i \cdot w *-y_i)X_{ij}$, where $X$ is the input matrix and $y$ is the input vector.
+To obtain a solution for linear regression using SGD, we instantiate the update function to a learned weight as $w_j = w_j  - \alpha(X_i \cdot w *-y_i)X_{ij}$, where $X$ is the input matrix and $y$ is the input vector.
 In this update function, the weights are adjusted element-wise by the error from the predicted and expected value at a rate determined by $\alpha$.
@@
 Logistic regression fits the logistic function $f(u) = \frac{1}{1+e^-u}$ to the input.
-The corresponding update function for mini-batched SGD for logistic regression is $ w = w - \frac{1}{|B|} \alpha X^{T}_{B} \times f(X_{B} \times w - Y_{B})$, where $f$ maps the predicted value into the binary output space.
+The corresponding update function for mini-batched SGD for logistic regression is $w = w - \frac{1}{|B|} \alpha X^{T}_{B} \times f(X_{B} \times w - Y_{B})$, where $f$ maps the predicted value into the binary output space.
 Mohassel and Zhang [8] proposed the following piecewise function as approximation for $f$:
@@
 It shows that the number of bits of precision needed to get good accuracy is highly dependent on the dataset.
-For MNIST, 13 bits was sufficient to match plaintext accuracy but 28 bits were needed for Arcene.
+For MNIST, 13 bits was sufficient to match plaintext accuracy, but 28 bits were needed for Arcene.
 The MNIST data contains only 784 features while there are 10,000 in the Arcene data, 3,000 of which are considered "probes" with no predictive power, which could explain the lower overall accuracy of \cite{MZ17}.
@@
 
-We found that the SPDZ outperformed Obliv-C even in a distributed machine setting by an order of magnitude and achieves comparable results to SecureML for SGD instantiations such as Linear and Logistic Regression.
-The accuracy findings were comparable for LDLT, Cholesky, and CGD results for SPDZ and Obliv-C, but we found that SPDZ outperformed Obliv-C on SGD.
-SPDZ was able to match SecureML results on SGD and Logistic Regression.
+SPDZ was able to achieve comparable accuracy for LDLT, Cholesky, and CGD when compared to Obliv-C with runtime faster by an order of magnitude on larger matrix sizes even in a distributed machine setting.
+SPDZ also achieved lower RMSE than Obliv-C using SGD.
+SPDZ was able to match accuracy and latency results for SecureML on SGD and Logistic Regression.
+This result is promising for SPDZ to be extended to more complex algorithms including neural networks with hidden layers.
 
@@
 
-\iffalse
+\newpage
 
-\small
+\section*{Supplementary Material}
 
-[1] Damg\aa rd, I., Keller, M., Larraia, E., Pastro, V., Scholl, P., \& Smart, N.
-P.
-(2013).
-Practical covertly secure MPC for dishonest majority?or: breaking the SPDZ limits.
-{\it European Symposium on Research in Computer Security}, pp.\ 1-18.
+\subsection{Three and four party runtimes}
 
-[2] Du, W., \& Atallah, M.
-J.
-(2001).
-Secure multi-party computation problems and their applications: a review and open problems.
-{\it Proceedings of the 2001 workshop on New security paradigms}, pp.\ 13-22.
-ACM.
+\begin{figure}[h]
+\centering
+  \caption{Run time results for 3 and 4 players in 32 bits. (Left) 3 Players. (Right) 4 Players.}
 
-[3] Friedman, J., Hastie, T., \& Tibshirani, R. (2001). The elements of statistical learning (Vol. 1, No. 10). New York, NY, USA:: Springer series in statistics.
+  \includegraphics[scale=0.6]{multi.jpg}
+  \label{fig:result4}
+\end{figure}
 
-[4] Gasc\'{o}n, A., Schoppmann, P., Balle, B., Raykova, M., Doerner, J., Zahur, S., \& Evans, D.
-(2017).
-Privacy-preserving distributed linear regression on high-dimensional data.
-{\it Proceedings on Privacy Enhancing Technologies, 2017(4)}, pp.\ 345-364.
+\subsection{EC2 runtimes}
 
-[5] Hazay, C., \& Lindell, Y. (2010). A Note on the Relation between the Definitions of Security for Semi-Honest and Malicious Adversaries. {\it IACR Cryptology ePrint Archive}, pp.\ 551.
+\begin{figure}[h]
+\centering
+  \caption{Run time results for 2 players deployed on EC2 instances averaged over 15 runs. (Left) 32 bits. (Right) 64 bits.}
+  \includegraphics[scale=0.7]{ec2.jpg}
+  \label{fig:result4}
+\end{figure}
 
-[6] Keller, M., Scholl, P., \& Smart, N.
-P.
-(2013).
-An architecture for practical actively secure MPC with dishonest majority.
-{\it Proceedings of the 2013 ACM SIGSAC conference on Computer \& communications security}, pp.\ 549-560.
-ACM.
+\subsection{Activation function runtime comparison}
 
-[7] Lindell, Y., \& Pinkas, B. (2009). A proof of security of Yao?s protocol for two-party computation. Journal of Cryptology, 22(2), 161-188.
+\begin{table}[h]
+\centering
+\label{my-label}
+\caption{Comparison of efficiency for the proposed activation function in [8] and Taylor Series approximation in SPDZ.}
+\begin{tabular}{@{}lcccc@{}}
+\toprule
+                        & \multicolumn{2}{c}{MNIST} & \multicolumn{2}{c}{Arcene} \\ \midrule
+                        & Compile      & Run        & Compile      & Run         \\
+New Activation Function & 26.01        & 18.8       & 220.14       & 224.47      \\
+2 Polynomial            & 27.79        & 19.28      & 222.85       & 239.98      \\
+5 Polynomial            & 30.32        & 19.8       & 225.33       & 241.59      \\
+7 Polynomial            & 31.99        & 20.65      & 228.33       & 239.11      \\
+10 Polynomial           & 34.24        & 20.71      & 230.81       & 240.79      \\ \bottomrule
+\end{tabular}
+\end{table}
 
-[8] Mohassel, P., \& Zhang, Y. (2017). SecureML: A system for scalable privacy-preserving machine learning. {\it 38th IEEE Symposium on Security and Privacy (SP)}, pp. 19-38. IEEE.
+\newpage
+\subsection{RMSE results}
 
-[9] Phung, D., Tseng, V.
-S., Webb, G.
-I., Ho, B., Ganji, M., \& Rashidi, L.
-(Eds.).
-(2018).
-Advances in Knowledge Discovery and Data Mining: 22nd Pacific-Asia Conference, PAKDD 2018, Melbourne, VIC, Australia, June 3-6, 2018, Proceedings (Vol.
-10938).
-Springer.
-
-[10] Zahur, S., \& Evans, D. (2015). Obliv-C: A Language for Extensible Data-Oblivious Computation. {\it IACR Cryptology ePrint Archive, 2015}, 1153.
-
-\fi
+\begin{table}[h]
+\centering
+\label{my-label}
+\caption{SGD Results for 9 open sourced UCI datasets.}
+\begin{tabular}{@{}llllll@{}}
+\toprule
+Dataset Name          & \multicolumn{1}{c}{SGD Plaintext} & \multicolumn{2}{c}{\begin{tabular}[c]{@{}c@{}}SPDZ (28 bit)\\ error / time\end{tabular}} & \multicolumn{2}{c}{\begin{tabular}[c]{@{}c@{}}SPDZ (13 bit)\\ error / time\end{tabular}} \\ \midrule
+Student Performance   & 0.11                              & 0.12 (+8.34\%)                                  & 2.174                                  & 0.12(+8.34\%)                                   & 2.616                                  \\
+Auto MPG              & 0.56                              & 0.68 (+21.4\%)                                  & 0.663                                  & 0.68(+21.4\%)                                   & 0.704                                  \\
+Communities and Crime & 0.06                              & 0.17 (+183\%)                                   & 10.352                                 & 0.19 (+216\%)                                   & 12.363                                 \\
+Wine Quality          & 0.18                              & 0.19 (+5.55\%)                                  & 0.809                                  & 0.19 (+5.55\%)                                  & 0.969                                  \\
+Bike Sharing          & 0.23                              & 0.24 (+4.34\%)                                  & 0.910                                  & 0.25 (+8.69\%)                                  & 1.097                                  \\
+Blog Feedback   & 0.04                              & 0.04 (+0.0\%)                                   & 9.455                                  & 0.04 (+0.0\%)                                   & 9.091                                  \\
+CT Slices   & 0.22                              & 0.22 (+0.0\%)                                   & 16.458                                  & 0.22 (+0.0\%)                                   & 16.295                                  \\
+Year Prediction MSD   & 0.06                              & 0.06 (+0.0\%)                                   & 7.146                                  & 0.06 (+0.0\%)                                   & 8.513                                  \\
+Gas Sensor Array      & 0.20                              & 0.36 (+20.0\%)                                  & 1.151                                  & 0.36 (+20.0\%)                                  & 1.559                                  \\ \bottomrule
+\end{tabular}\\
+\end{table}
 
 \end{document}
-
-%Can add table back in - I am describing results instead. 
-%\begin{table}[h!]
-%\caption{Runtime results against }
-%\centering
-%\begin{tabular}{@{}llcl@{}}
-%\toprule
-%\multicolumn{1}{c}{Datasets}                                           & \multicolumn{3}{c}{Runtime (seconds)}                                      \\ \midrule
-%CGD (64 bit)                                                           & \multicolumn{1}{c}{Ours} & Ours (EC2) & \multicolumn{1}{c}{Obliv-C}        \\ \midrule
-%d=100                                                                  & 144                      & 243        & \textgreater $10^3$ \\ \midrule
-%d=500                                                                  & 527                      & 2701       & \textgreater $10^4$ \\ \midrule
-%\multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}SGD: \\ MNIST\end{tabular}} & \multicolumn{1}{c}{Ours} & \multicolumn{2}{c}{SecureML}                    \\ \cmidrule(l){2-4} 
-%                                                                       & 10200                    & \multicolumn{2}{c}{10330}                       \\ \bottomrule
-%\end{tabular}
-%\end{table}
-
-%\section{Conclusion}
-
-%Through this work, we demonstrated that SPDZ, a malicious security framework, can match the efficiency and accuracy other protocols of semi-honest through experiments on real and synthetic data.
-This investigation into the applicability of secure MPC on real world data demonstrates its potential change the way data is used in a variety of applications in today\textquotesingle s world.

### 2018-10-16 paper/nips_format.tex
edits

@@
 Most of our experiments were ran using machines on the same local area network where there is no network latency.
-We performed tests where both parties were deployed on separate Amazon EC2 m4.large instances.
-We also ran experiment with up to four parties.
+We performed tests where both parties were deployed on separate Amazon EC2 m4.large instances (see Figure~\ref{fig:result4}).
+We also ran experiment with up to four parties (see Figure~\ref{fig:result3}).
 
@@
   \includegraphics[scale=0.6]{multi.jpg}
-  \label{fig:result4}
+  \label{fig:result3}
 \end{figure}
@@
 \end{tabular}
+\label{tbl2}
 \end{table}
@@
 \end{tabular}\\
+\label{tbl3}
 \end{table}

### 2018-10-16 paper/nips_format.tex
minor edits, added sentences to supplements

@@
 \begin{abstract}
-Secure Multi-Party Computation (MPC) is an area of cryptography that enables the computation on sensitive data from multiple sources while maintaining privacy guarantees.
-However, theoretical MPC protocols often do not scale efficiently to real-world data. This project investigates the efficiency of of the SPDZ framework~\cite{SPDZ12}, which provides an implementation
+Secure Multi-Party Computation (MPC) is an area of cryptography that enables computation on sensitive data from multiple sources while maintaining privacy guarantees.
+However, theoretical MPC protocols often do not scale efficiently to real-world data. This project investigates the efficiency of the SPDZ framework~\cite{SPDZ12}, which provides an implementation
 of an MPC protocol with malicious security, in the context of popular machine learning (ML) algorithms. In particular, we chose applications such as linear regression and logistic regression,
@@
 The problem is that openly sharing this data for prediction or model-building purposes is against modern day privacy laws as it would leak private individual data.
-This is one of many examples, which could benefit from MPC which would allow the parties to evaluate the output of the analysis without revealing more about the private inputs.
+This is one of many real-world applications that could benefit from MPC, which allows parties to evaluate the output of the analysis without revealing more about the private inputs.
 %This application of secure MPC is one that would be extremely valuable in today\textquotesingle s world.
@@
 
-Two recent works~\cite{GSB0DZE17, MZ17} propose efficient implementation for several central building blocks for machine learning such as conjugate gradient decent (CGD) and stochastic gradient descent (SGD) as well as applications using them such as linear and logistic regression.
+Two recent works~\cite{GSB0DZE17, MZ17} propose efficient implementation of several central machine learning building blocks including conjugate gradient decent (CGD) and stochastic gradient descent (SGD) as well as their applications in linear and logistic regression.
 The work of Gascon et al.~\cite{GSB0DZE17} uses the framework for semi-honest computation Obliv-C~\cite{cryptoeprint:2015:1153} and proposes
@@
 return for a small accuracy loss.
-They further propose a modification of CGD that has stable behavior using fixed-point arithmetic as emulating floating point with the underlying MPC representation introduces substantial
+They further propose a modification of CGD that has stable behavior using fixed-point arithmetic because emulating floating point with the underlying MPC representation introduces substantial
 efficiency overhead.
 
-The work of Mohassel and Zhang~\cite{MZ17} considers stochastic gradient descent as a method for linear regression and logistic regression with the incorporation of different activation functions.
+The work of Mohassel and Zhang~\cite{MZ17} utilizes stochastic gradient descent as a method for linear regression and logistic regression with the incorporation of different activation functions.
 The authors consider
-arithmetic representation for the computation and propose new secure computation techniques for matrix computation, which generalize the approach for generating multiplicative triples in a preprocessing step.
+arithmetic representation for the computation and propose new secure computation techniques for matrix computation, which generalizes the approach for generating multiplicative triples in a preprocessing step.
  Similarly to the work of Gascon et al.~\cite{GSB0DZE17}, this paper considers techniques for approximation that save computation.
@@
 The difference between LDLT and Cholesky is that Cholesky requires a square root. The representation of square root computation as an arithmetic circuit used in the MPC computation in SPDZ introduces 
-considerable overhear. That is why we used the iterative Newton method as a way of approximating the square root computation. It computes $x_i$, where $x_{i}^2 = S$, with repetition of the following
-update function $x_{n+1} =1/2 (x_{n} + S/x_{n})$.
+considerable overhead. That is why we used the iterative Newton method as a way of approximating the square root computation. It computes $x_i$, where $x_{i}^2 = S$, with repetition of the following
+update function $x_{n+1} =\frac{1}{2}(x_{n} + S/x_{n})$.
 
@@
 
-To obtain a solution for linear regression using SGD, we instantiate the update function to a learned weight as $w_j = w_j  - \alpha(X_i \cdot w *-y_i)X_{ij}$, where $X$ is the input matrix and $y$ is the input vector.
+To obtain a solution for linear regression using SGD, we instantiate the update function of a learned weight as $w_j = w_j  - \alpha(X_i \cdot w *-y_i)X_{ij}$, where $X$ is the input matrix and $y$ is the input vector.
 In this update function, the weights are adjusted element-wise by the error from the predicted and expected value at a rate determined by $\alpha$.
@@
 
-In this section we present empirical results for our SPDZ implementations evaluated with real and synthetic databases.
+In this section we present empirical results for our SPDZ implementations evaluated with real and synthetically generated databases.
 We compare the five different algorithms in terms of accuracy and run time for various parameters.
@@
 
+In addition to the 2 player case, we also ran experiments with 3 and 4 players on the local network, where the input was vertically partitioned between parties.
+
 \begin{figure}[h]
 \centering
-  \caption{Run time results for 3 and 4 players in 32 bits. (Left) 3 Players. (Right) 4 Players.}
+  \caption{Run time results for 3 and 4 players in 32 bits. (Left) 3 Players. (Right) 4 Players. }
 
@@
 \subsection{EC2 runtimes}
+
+We deployed 2 AWS EC2 Instances that ran the SPDZ protocol.
+The EC2 runtime results are comparable to the local network results for smaller matrix sizes but increases at a faster rate for larger matrix sizes.
 
@@
 
+To determine the difference in efficiency in the two activation functions, the new one proposed by ~\cite{MZ17} and the other method which they claimed to be inefficient, we ran results for SPDZ on two different datasets for a few different variations of the activation functions.
+The compile and run times correspond to the offline and online phase for SPDZ respectively.
+
 \begin{table}[h]
@@
 \label{my-label}
-\caption{Comparison of efficiency for the proposed activation function in [8] and Taylor Series approximation in SPDZ.}
+\caption{Comparison of efficiency for the proposed activation function in ~\cite{MZ17} and Taylor Series approximation in SPDZ.}
 \begin{tabular}{@{}lcccc@{}}
@@
 
+SGD was also evaluated on 9 different datasets selected from the UCI repository.
+Further details about the specific datasets, including the regularization parameter used, are detailed in ~\cite{GSB0DZE17}.
+The dimensions of the problems ranged from 7 to 384 with the number of examples ranging from over 200 to almost 3 million.
+
 \begin{table}[h]
@@
 \caption{SGD Results for 9 open sourced UCI datasets.}
+\newcommand*{\MyIndent}{\hspace*{1.2cm}}%
 \begin{tabular}{@{}llllll@{}}
 \toprule
-Dataset Name          & \multicolumn{1}{c}{SGD Plaintext} & \multicolumn{2}{c}{\begin{tabular}[c]{@{}c@{}}SPDZ (28 bit)\\ error / time\end{tabular}} & \multicolumn{2}{c}{\begin{tabular}[c]{@{}c@{}}SPDZ (13 bit)\\ error / time\end{tabular}} \\ \midrule
+Dataset Name          & \multicolumn{1}{c}{SGD Plaintext} & \multicolumn{2}{c}{\begin{tabular}[c]{@{}c@{}}SPDZ (28 bit)\\ Error \MyIndent Time\end{tabular}} & \multicolumn{2}{c}{\begin{tabular}[c]{@{}c@{}}SPDZ (13 bit)\\ Error \MyIndent Time\end{tabular}} \\ \midrule
 Student Performance   & 0.11                              & 0.12 (+8.34\%)                                  & 2.174                                  & 0.12(+8.34\%)                                   & 2.616                                  \\
